---
title: "Rarefied Diversity Analysis"
author: "Marian L. Schmidt"
date: "February 14, 2017"
output:
  html_document:
    code_folding: show
    highlight: haddock
    keep_md: yes
    theme: united
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
---
<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

```{r setup, include=FALSE}
# For width of code chunks and scroll bar 
options(width=250)

knitr::opts_chunk$set(eval = TRUE, 
                      echo = TRUE, 
                      cache = TRUE,
                      include = TRUE,
                      warning = TRUE,
                      collapse = FALSE,
                      message = FALSE,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = TRUE,
                      fig.path="Rarefied_Figures/",  # Set the figure options
                      fig.align = "center", 
                      fig.width = 7,
                      fig.height = 7)

```


```{r load-libraries, message = FALSE, warning = FALSE}
library(ggplot2)
library(devtools)
library(phyloseq)
library(tidyr)
library(dplyr)
library(cowplot)
library(picante) # Will also include ape and vegan 
library(car) # For residual analysis
library(sandwich) # for vcovHC function in post-hoc test
library(MASS) # For studres in plot_residuals function
source("Muskegon_functions.R")
source("set_colors.R")
```


# Rarefied Alpha Diversity Analysis
## Remove samples with too few reads
```{r load-data, eval = TRUE}
# Loads a phyloseq object named otu_merged_musk_pruned)
load("../data/otu_merged_musk_pruned.RData")
# The name of the phyloseq object is: 
otu_merged_musk_pruned 

# Productivity measurements are reliable only up to 1 decimal
df1 <- sample_data(otu_merged_musk_pruned) %>% 
  dplyr::mutate(tot_bacprod = round(tot_bacprod, digits = 1),
                SD_tot_bacprod = round(SD_tot_bacprod, digits = 1),
                frac_bacprod = round(frac_bacprod, digits = 1),
                SD_frac_bacprod = round(SD_frac_bacprod, digits = 1))
row.names(df1) = df1$norep_filter_name


# Remove tree because it's too computationally intensive
otu_merged_musk_pruned <- merge_phyloseq(tax_table(otu_merged_musk_pruned), 
                                         sample_data(df1), 
                                         otu_table(otu_merged_musk_pruned))
otu_merged_musk_pruned

# Remove MOTHJ715 and MBRHP715
otu_merged_musk_pruned_noMOTHJ715_MBRHP715 <- subset_samples(otu_merged_musk_pruned, 
                                                             norep_filter_name != "MOTHJ715" & 
                                                               norep_filter_name != "MBRHP715")
otu_merged_musk_pruned_noMOTHJ715_MBRHP715
```


```{r seq-read-depth, eval = TRUE}
# Check the sequencing depth of each sample by 
sums_otu <- data.frame(rowSums(otu_table(otu_merged_musk_pruned_noMOTHJ715_MBRHP715))) 
colnames(sums_otu) <- "Sample_TotalSeqs"
sums_otu$names <- row.names(sums_otu)
sums_otu <- arrange(sums_otu, Sample_TotalSeqs) 
sums_otu <- make_metadata_norep(sums_otu)

## Add total sequences to metadata frame 
metdf <- sample_data(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
sums_otu$norep_filter_name <- sums_otu$names
sums_otu2 <- dplyr::select(sums_otu, norep_filter_name, Sample_TotalSeqs)
metdf_num2 <- left_join(metdf, sums_otu2, by = "norep_filter_name") %>%
  dplyr::select(-one_of("D0", "D0_chao", "D1", "D2", "D0_SD", "D1_sd", "D0_chao_sd"))
row.names(metdf_num2) <- metdf_num2$norep_filter_name
# Rename the sample data 
sample_data(otu_merged_musk_pruned_noMOTHJ715_MBRHP715) <- metdf_num2
```

## Metadata Frames for Subsets
```{r metadata-frames}
### PREPARE DATA FRAMES FOR PHENOFLOW ALPHA DIVERSITY ANALYSIS
free_meta_data <- filter(metdf_num2, fraction %in% c("Free", "WholeFree") & norep_filter_name != "MOTHJ715" & limnion == "Top")
part_meta_data <- filter(metdf_num2, fraction %in% c("Particle", "WholePart") & norep_filter_name != "MOTHJ715" & limnion == "Top")
nosed_meta_data <- filter(metdf_num2, fraction != "Sediment" & limnion == "Top")

# Only "True Free Living and Particle" From 20um prefiltered samples 
free_only <- filter(free_meta_data, fraction == "Free")
part_only <- filter(part_meta_data, fraction == "Particle")

## 2015 specific samples that have NOT been prefiltered (whole water)
wholefree_only <- filter(free_meta_data, fraction == "WholeFree")
wholepart_only <- filter(part_meta_data, fraction == "WholePart")
```


#### Rarefy Read Depth Analysis ran on February 14th, 2016
```{r calculate-alphadiv, eval = FALSE, echo = FALSE}
 # Initialize matrices to store richness and evenness estimates
# nsamp <- nsamples(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
# min_lib <- min(sample_sums(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)) - 1
# 
# otu_richness <- matrix(nrow = nsamp, ncol = 100)
# row.names(otu_richness) <- sample_names(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
# 
# otu_evenness <- matrix(nrow = nsamp, ncol = 100)
# row.names(otu_evenness) <- sample_names(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
# 
# otu_shannon <- matrix(nrow = nsamp, ncol = 100)
# row.names(otu_shannon) <- sample_names(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
## 
 ## It is always important to set a seed when you subsample so your result is replicable
 #set.seed(777)

# for (i in 1:100) {
#  # Subsample
#  r <- rarefy_even_depth(otu_merged_musk_pruned_noMOTHJ715_MBRHP715, sample.size = min_lib, verbose = FALSE, replace = TRUE)
# 
#  # Calculate richness
#  rich <- as.numeric(as.matrix(estimate_richness(r, measures = "Observed")))
#  otu_richness[ ,i] <- rich
# 
#  # Calculate evenness
#  even <- as.numeric(as.matrix(estimate_richness(r, measures = "InvSimpson")))
#  otu_evenness[ ,i] <- even
# 
#  # Calculate evenness
#  shannon <- as.numeric(as.matrix(estimate_richness(r, measures = "Shannon")))
#  otu_shannon[ ,i] <- shannon
# 
# }

#  write.table(otu_evenness, "../data/metadata/otu_evenness100_rarefy2895", row.names = TRUE)
#  write.table(otu_richness, "../data/metadata/otu_richness100_rarefy2895", row.names = TRUE)
#  write.table(otu_shannon, "../data/metadata/otu_shannon100_rarefy2895", row.names = TRUE)
```

# Fraction Diversity-Production Analysis 
```{r vean-otu-alphadiv-calc, eval = TRUE, fig.width = 10, fig.height = 8, warning = FALSE}
# Load values
nsamp <- nsamples(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
min_lib <- min(sample_sums(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)) - 1
min_lib

# Read in the files 
otu_richness <- read.table("../data/metadata/otu_richness100_rarefy2895",  header = TRUE)
otu_evenness <- read.table("../data/metadata/otu_evenness100_rarefy2895", header = TRUE)
otu_shannon <- read.table("../data/metadata/otu_shannon100_rarefy2895", header = TRUE)

# Create a new dataframe to hold the means and standard deviations of richness estimates
norep_filter_name <- row.names(otu_richness)
mean <- apply(otu_richness, 1, mean)
sd <- apply(otu_richness, 1, sd)
measure <- rep("Richness", nsamp)
otu_rich_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Create a new dataframe to hold the means and standard deviations of evenness estimates
norep_filter_name <- row.names(otu_evenness)
mean <- apply(otu_evenness, 1, mean)
sd <- apply(otu_evenness, 1, sd)
measure <- rep("Inverse_Simpson", nsamp)
otu_even_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Create a new dataframe to hold the means and standard deviations of shannon entropy estimates
norep_filter_name <- row.names(otu_shannon)
mean <- apply(otu_shannon, 1, mean)
sd <- apply(otu_shannon, 1, sd)
measure <- rep("Shannon_Entropy", nsamp)
otu_shan_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Create a new dataframe to hold the means and standard deviations of simpsons evenness estimates
norep_filter_name <- row.names(otu_evenness)
mean <- apply(otu_evenness, 1, mean)
sd <- apply(otu_evenness, 1, sd)
measure <- rep("Inverse_Simpson", nsamp)
otu_simpseven_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Calculate Simpson's Evenness into new df called "simps_evenness"
otu_simps_evenness <- inner_join(otu_rich_stats, otu_even_stats, by = "norep_filter_name") %>%
  mutate(mean = mean.y/mean.x,
         sd = sd(mean),
         measure = "Simpsons_Evenness") %>%
  dplyr::select(norep_filter_name, mean, sd, measure)

# Combine alpha diversity into one dataframe 
otu_alpha <- rbind(otu_rich_stats, otu_even_stats, otu_simps_evenness, otu_shan_stats)
s <- data.frame(sample_data(otu_merged_musk_pruned_noMOTHJ715_MBRHP715))
otu_alphadiv <- merge(otu_alpha, s, by = "norep_filter_name") %>%
  filter(project == "Muskegon_Lake" & limnion == "Top" & fraction != "Sediment") %>%
  mutate(fraction = factor(fraction, levels = c("WholePart", "Particle", "WholeFree", "Free")),
         lakesite = factor(lakesite,  levels = c("MOT", "MDP", "MBR", "MIN")),
         measure = factor(measure, levels = c("Richness", "Simpsons_Evenness", "Shannon_Entropy", "Inverse_Simpson"))) %>%
  # Calculate the per cell production rates 
  mutate(fraction_bac_abund = as.numeric(fraction_bac_abund),
         fracprod_per_cell = frac_bacprod/(1000*fraction_bac_abund),
         fracprod_per_cell_noinf = ifelse(fracprod_per_cell == Inf, NA, fracprod_per_cell))
```


## Subset Diversity Data 
```{r subset-diversity, dependson = c("load-data","vean-otu-alphadiv-calc")}
######################################################### RICHNESS
# Subset only richness data 
ML_otu_rich_stats <- filter(otu_alphadiv, measure == "Richness" & 
                              project == "Muskegon_Lake" & 
                              fraction %in% c("WholePart", "WholeFree") & year == "2015")

######################################################### SHANNON ENTROPY
# Subset only Shannon_Entropy data 
ML_otu_shannon_stats <- filter(otu_alphadiv, 
                               measure == "Shannon_Entropy" & 
                                 project == "Muskegon_Lake" & 
                                 fraction %in% c("WholePart", "WholeFree") & 
                                 year == "2015")

######################################################### INVERSE SIMPSON
# Subset only Inverse_Simpson data 
ML_otu_invsimps_stats <- filter(otu_alphadiv, 
                                measure == "Inverse_Simpson" & 
                                  project == "Muskegon_Lake" & 
                                  fraction %in% c("WholePart", "WholeFree") & 
                                  year == "2015")

######################################################### SIMPSON'S EVENNESS
# Subset only Simpsons_Evenness data 
ML_otu_simpseven_stats <- filter(otu_alphadiv, 
                                 measure == "Simpsons_Evenness" & 
                                   project == "Muskegon_Lake" & 
                                   fraction %in% c("WholePart", "WholeFree") & 
                                   year == "2015")
```



# Cell Count and Production Rates 
```{r boxplot-cellcount-prod, eval = TRUE, fig.width = 10, fig.height = 4, warning = FALSE, dependson="subset-diversity"}
######################################################### Fraction ABUNDANCe 
frac_abund_wilcox <- wilcox.test(log10(as.numeric(fraction_bac_abund)) ~ fraction, 
             data = filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Richness"))
frac_abund_wilcox

filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Richness") %>%
  group_by(fraction) %>%
  summarize(mean(as.numeric(fraction_bac_abund)))

# Make a data frame to draw significance line in boxplot (visually calculated)
dat1 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(6.45,6.5,6.5,6.45)) # WholePart vs WholeFree

poster_a <- ggplot(filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & 
                            measure == "Richness" & norep_filter_name != "MOTEJ515"), 
       aes(y = log10(as.numeric(fraction_bac_abund)), x = fraction)) +
  scale_color_manual(values = fraction_colors) + 
  scale_fill_manual(values = fraction_colors) +
  geom_jitter(size = 3, aes(color = fraction, fill = fraction), width = 0.2) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  ylab("Log10(Bacterial Cells/mL)") +
  scale_x_discrete(breaks=c("WholePart", "WholeFree"),
                      labels=c("Particle-\nAssociated", "Free-\nLiving")) + 
  #####  WHOLE PARTICLE VS WHOLE FREE CELL ABUNDANCES
  geom_path(data = dat1, aes(x = a, y = b), linetype = 1, color = "gray40") +
  annotate("text", x=1.5, y=6.5, fontface = "bold",  size = 3.5, color = "gray40",
           label= paste("***\np =", round(frac_abund_wilcox$p.value, digits = 6))) +
  theme(legend.position = "none",
        axis.title.x = element_blank())



######################################################### TOTAL PRODUCTION 
totprod_wilcox <- wilcox.test(frac_bacprod ~ fraction, 
             data = filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Richness"))
totprod_wilcox

filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Richness") %>%
  group_by(fraction) %>%
  summarize(mean(frac_bacprod))

# Make a data frame to draw significance line in boxplot (visually calculated)
dat2 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(67,68,68,67)) # WholePart vs WholeFree


poster_b <- ggplot(filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & measure == "Richness"), 
       aes(y = frac_bacprod, x = fraction)) + 
  scale_color_manual(values = fraction_colors) + 
  scale_fill_manual(values = fraction_colors) +
  geom_jitter(size = 3, aes(color = fraction, fill = fraction), width = 0.2) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  ylab("Secondary Production (μgC/L/hr)") +
  scale_x_discrete(breaks=c("WholePart", "WholeFree"),
                    labels=c("Particle-\nAssociated", "Free-\nLiving")) + 
  #####  WHOLE PARTICLE VS WHOLE FREE TOTAL PRODUCTION 
  geom_path(data = dat2, aes(x = a, y = b), linetype = 1, color = "gray40") +
  annotate("text", x=1.5, y=68, fontface = "bold",  size = 3.5, color = "gray40",
           label= paste("***\np =", round(totprod_wilcox$p.value, digits = 3))) +
  theme(legend.position = "none",
        axis.title.x = element_blank())



######################################################### TOTAL PRODUCTION 
percellprod_wilcox <- wilcox.test(log10(fracprod_per_cell) ~ fraction, 
             data = filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & 
                            measure == "Richness" &
                            norep_filter_name != "MOTEJ515"))
percellprod_wilcox

filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Richness" &
       norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515") %>%
  group_by(fraction) %>%
  summarize(mean(fracprod_per_cell))


# Make a data frame to draw significance line in boxplot (visually calculated)
dat3 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(-5.05,-5,-5,-5.05)) # WholePart vs WholeFree


poster_c <- ggplot(filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & 
                            measure == "Richness" &
                            norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515"), 
       aes(y = log10(fracprod_per_cell), x = fraction)) +
  scale_color_manual(values = fraction_colors) + 
  scale_fill_manual(values = fraction_colors) +
  geom_jitter(size = 3, aes(color = fraction, fill = fraction), width = 0.2) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  ylim(c(-8.5, -5)) + 
  ylab("log10(Production/cell) (μgC/cell/hr)") +
  scale_x_discrete(breaks=c("WholePart", "WholeFree"),
                    labels=c("Particle-\nAssociated", "Free-\nLiving")) + 
  #####  WHOLE PARTICLE VS WHOLE FREE PER CELL PRODUCTION 
  geom_path(data = dat3, aes(x = a, y = b), linetype = 1, color = "gray40") +
  annotate("text", x=1.5, y=-5, fontface = "bold",  size = 3.5, color = "gray40",
           label= paste("***\np =", round(percellprod_wilcox$p.value, digits = 5))) +
  theme(legend.position = "none",
        axis.title.x = element_blank())

plot_grid(poster_a, poster_b, poster_c,
          labels = c("A", "B", "C"),
          ncol = 3)
```



# Diversity Comparison
```{r boxplot-diversity-comparison, fig.width = 10, fig.height = 4}
######################################################### RICHNESS 
rich_wilcox <- wilcox.test(mean ~ fraction, 
             data = filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Richness"))

filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Richness") %>%
  group_by(fraction) %>%
  summarize(mean(mean),  median(mean))

# Make a data frame to draw significance line in boxplot (visually calculated)
nums1 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(790,800,800,790)) # WholePart vs WholeFree

poster_rich1 <-  ggplot(filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & measure == "Richness"), 
       aes(y = mean, x = fraction)) +
  ylab("Observed Richness") + 
  #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd, color = fraction), alpha = 0.7) + # X-axis errorbars
  geom_jitter(size = 3, aes(color = fraction, fill = fraction), width = 0.2) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  geom_path(data = nums1, aes(x = a, y = b), linetype = 1, color = "gray40") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(180, 810), breaks = c(200, 400, 600, 800)) 
     
     
poster_rich <- poster_rich1 + 
  annotate("text", x=1.5, y=800, fontface = "bold",  size = 3.5, color = "gray40",
           label= paste("***\np =", round(rich_wilcox$p.value, digits = 3))) +
  theme(legend.position = c(0.7, 0.75), 
        axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.spacing.y = unit(-0.5, "cm"),
        legend.spacing.x = unit(-0.3, "cm"),
        legend.box = "horizontal") +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free", "Particle")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free", "Particle")) +
  geom_jitter(size = 3, aes(color = fraction, fill = fraction, shape = lakesite), width = 0.2) + 
  scale_shape_discrete(guide = guide_legend(reverse=TRUE))



######################################################### SHANNON ENTROPY  
shannon_wilcox <- wilcox.test(mean ~ fraction, 
             data = filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Shannon_Entropy"))

filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Shannon_Entropy") %>%
  group_by(fraction) %>%
  summarize(mean(mean),  median(mean))

# Make a data frame to draw significance line in boxplot (visually calculated)
nums2 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(5.75,5.8,5.8,5.75)) # WholePart vs WholeFree

poster_shannon1 <- ggplot(filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & measure == "Shannon_Entropy"), 
       aes(y = mean, x = fraction)) +  
  ylab("Shannon Entropy") +   
  #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.7) + # X-axis errorbars
  geom_jitter(size = 3, aes(color = fraction, fill = fraction), width = 0.2) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  geom_path(data = nums2, aes(x = a, y = b), linetype = 1, color = "gray40") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(3.4, 5.85), breaks = c(3.5, 4, 4.5, 5, 5.5))  

  
  
poster_shannon <- poster_shannon1 + 
  annotate("text", x=1.5, y=5.8, fontface = "bold",  size = 3.5, color = "gray40",
           label= paste("***\np =", round(shannon_wilcox$p.value, digits = 3))) +
  theme(legend.position = "none", axis.title.x = element_blank())


#########################################################  INVERSE SIMPSON 
simpson_wilcox <- wilcox.test(mean ~ fraction, 
             data = filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & 
                            measure == "Inverse_Simpson" &
                            norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515"))
simpson_wilcox

filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Inverse_Simpson" &
       norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515") %>%
  group_by(fraction) %>%
  summarize(mean(mean), median(mean))

nums3 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(88,89, 89, 88)) # WholePart vs WholeFree


poster_invsimps1 <-   ggplot(filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & 
                            measure == "Inverse_Simpson" &
                            norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515"), 
       aes(y = mean, x = fraction)) +
  ylab("Inverse Simpson") +   
  #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.7) + # X-axis errorbars
  geom_jitter(size = 3, aes(color = fraction, fill = fraction), width = 0.2) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  geom_path(data = nums3, aes(x = a, y = b), linetype = "dotted", color = "gray40") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0,95), breaks = c(0, 20, 40, 60, 80), expand = c(0,0))  


poster_invsimps <- poster_invsimps1 +
  annotate("text", x=1.5, y=89, fontface = "bold",  size = 3.5, color = "gray40",
           label= paste("NS\np =", round(simpson_wilcox$p.value, digits = 2))) +
  theme(legend.position = "none", axis.title.x = element_blank())

#########################################################  SIMPSON'S EVENNESS
simpseven_wilcox <- wilcox.test(mean ~ fraction, 
             data = filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & 
                            measure == "Simpsons_Evenness" &
                            norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515"))
simpseven_wilcox

filter(otu_alphadiv, fraction %in% c("WholePart", "WholeFree") & measure == "Simpsons_Evenness" &
       norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515") %>%
  group_by(fraction) %>%
  summarize(mean(mean), median(mean))

nums3 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(0.145, 0.15, 0.15, 0.145)) # WholePart vs WholeFree


poster_simpseven1 <-   ggplot(filter(otu_alphadiv, 
                          fraction %in% c("WholePart", "WholeFree") & 
                            measure == "Simpsons_Evenness" &
                            norep_filter_name != "MOTEJ515" & norep_filter_name != "MOTEP515"), 
       aes(y = mean, x = fraction)) +
  ylab("Simpson's Evenness") +   
  #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.7) + # X-axis errorbars
  geom_jitter(size = 3, aes(color = fraction, fill = fraction), width = 0.2) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  geom_path(data = nums3, aes(x = a, y = b), linetype = "dotted", color = "gray40") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0.03,0.151), breaks = c(0.05, 0.1, 0.15))  

  
poster_simpseven <- poster_simpseven1 +
  annotate("text", x=1.5, y=0.15, fontface = "bold",  size = 3.5, color = "gray40",
           label= paste("NS\np =", round(simpseven_wilcox$p.value, digits = 2))) +
  theme(legend.position = "none", axis.title.x = element_blank())

plot_grid(poster_rich, poster_shannon, poster_invsimps, poster_simpseven,
          labels = c("A", "B", "C", "D"),
          ncol = 4)

```


# Diversity vs Fraction Production 
```{r plot-vegan-fracprod, dependson = "subset-diversity", fig.width = 10, fig.height = 10, warnings = FALSE}
######################################################### RICHNESS
# Free-Living Richness vs fractional production 
freeprod_ML_otu_rich <- lm(frac_bacprod ~ mean, data = filter(ML_otu_rich_stats, fraction == "WholeFree"))
summary(freeprod_ML_otu_rich) 

# Particle-Associated Richness vs fractional production 
partprod_MLotu_rich <- lm(frac_bacprod ~ mean, data = filter(ML_otu_rich_stats, fraction == "WholePart"))
summary(partprod_MLotu_rich) 

# Both fractions together
summary(lm(frac_bacprod ~ mean, data = ML_otu_rich_stats))
 
# Plot 
otu_rich_vegan <-  ggplot(ML_otu_rich_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + 
  geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd), alpha = 0.7) + # X-axis errorbars
  # Y-axis errorbars
  geom_errorbar(aes(ymin = frac_bacprod - SD_frac_bacprod, ymax = frac_bacprod + SD_frac_bacprod),  alpha = 0.5) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), 
                     limits = c("WholePart", "WholeFree"),
                     breaks=c("WholePart", "WholeFree"),
                     labels=c("Particle", "Free")) + 
  ylab("Secondary Production (μgC/L/hr)") + xlab("Observed Richness") +
  geom_smooth(data=subset(ML_otu_rich_stats, fraction == "WholePart"), method='lm') + 
  scale_x_continuous(limits = c(180, 810), breaks = c(200, 400, 600, 800)) + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 250, y=55, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_rich)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_rich)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 650, y=3, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_rich)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_rich)$coefficients[,4][2]), digits = 4)));


######################################################### SHANNON ENTROPY
# Free-Living Shannon vs fractional production 
freeprod_ML_otu_shannon <- lm(frac_bacprod ~ mean, data = filter(ML_otu_shannon_stats, fraction == "WholeFree"))
summary(freeprod_ML_otu_shannon)

# Particle-Associated Shannon vs fractional production 
partprod_MLotu_shannon <- lm(frac_bacprod ~ mean, data = filter(ML_otu_shannon_stats, fraction == "WholePart"))
summary(partprod_MLotu_shannon)

# Both fractions together
summary(lm(frac_bacprod ~ mean, data = ML_otu_shannon_stats))

# Plot 
otu_shannon_vegan <- ggplot(ML_otu_shannon_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + 
  geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd), alpha = 0.7) + # X-axis errorbars
  # Y-axis errorbars
  geom_errorbar(aes(ymin = frac_bacprod - SD_frac_bacprod, ymax = frac_bacprod + SD_frac_bacprod),  alpha = 0.5) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), 
                     limits = c("WholePart", "WholeFree"),
                     breaks=c("WholeFree", "WholePart"),
                     labels=c("Free-Living", "Particle-Associated")) + 
  ylab("Secondary Production (μgC/L/hr)") + xlab("Shannon Entropy") +
  scale_x_continuous(limits = c(3.4, 5.85), breaks = c(3.5, 4, 4.5, 5, 5.5)) + 
  geom_smooth(data=subset(ML_otu_shannon_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 3.75, y=55, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_shannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_shannon)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 5.35, y=3, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_shannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_shannon)$coefficients[,4][2]), digits = 4))); 


######################################################### INVERSE SIMPSON
# Free-Living Inverse Simpson vs fractional production 
freeprod_ML_otu_invsimps <- lm(frac_bacprod ~ mean, data = filter(ML_otu_invsimps_stats, fraction == "WholeFree"))
summary(freeprod_ML_otu_invsimps)

# Particle-Associated Inverse Simpson vs fractional production 
partprod_MLotu_invsimps <- lm(frac_bacprod ~ mean, data = filter(ML_otu_invsimps_stats, fraction == "WholePart"))
summary(partprod_MLotu_invsimps)

# Both fractions together
summary(lm(frac_bacprod ~ mean, data = ML_otu_invsimps_stats))

# Plot Simpson's Evenness
otu_invsimps_vegan <- ggplot(ML_otu_invsimps_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) +  
  geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd), alpha = 0.7) + # X-axis errorbars
  # Y-axis errorbars
  geom_errorbar(aes(ymin = frac_bacprod - SD_frac_bacprod, ymax = frac_bacprod + SD_frac_bacprod),  alpha = 0.5) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), 
                     limits = c("WholePart", "WholeFree"),
                     breaks=c("WholeFree", "WholePart"),
                     labels=c("Free-Living", "Particle-Associated")) + 
  scale_x_continuous(limits = c(0,95), breaks = c(0, 20, 40, 60, 80), expand = c(0,0)) + 
  ylab("Secondary Production (μgC/L/hr)") + xlab("Inverse Simpson") +
  geom_smooth(data=subset(ML_otu_invsimps_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.85,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 15, y=55, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_invsimps)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_invsimps)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 63, y=3, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_invsimps)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_invsimps)$coefficients[,4][2]), digits = 4))); 



######################################################### SIMPSON'S EVENNESS
# Free-Living Simpson's Evenness vs fractional production 
freeprod_ML_otu_simpseven <- lm(frac_bacprod ~ mean, data = filter(ML_otu_simpseven_stats, fraction == "WholeFree"))
summary(freeprod_ML_otu_simpseven)

# Particle-Associated Simpson's Evenness vs fractional production 
partprod_MLotu_simpseven <- lm(frac_bacprod ~ mean, data = filter(ML_otu_simpseven_stats, fraction == "WholePart"))
summary(partprod_MLotu_simpseven)

# Both fractions together
summary(lm(frac_bacprod ~ mean, data = ML_otu_simpseven_stats))

# Plot 
otu_simpseven_vegan <- ggplot(ML_otu_simpseven_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) +  
  geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd), alpha = 0.7) + # X-axis errorbars
  # Y-axis errorbars
  geom_errorbar(aes(ymin = frac_bacprod - SD_frac_bacprod, ymax = frac_bacprod + SD_frac_bacprod),  alpha = 0.5) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), 
                     limits = c("WholePart", "WholeFree"),
                     breaks=c("WholeFree", "WholePart"),
                     labels=c("Free-Living", "Particle-Associated")) + 
  scale_x_continuous(limits = c(0.03,0.151), breaks = c(0.05, 0.1, 0.15))  +
  ylab("Secondary Production (μgC/L/hr)") + xlab("Simpson's Evenness") +
  geom_smooth(data=subset(ML_otu_simpseven_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 0.05, y=55, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_simpseven)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_simpseven)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 0.125, y=3, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_simpseven)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_simpseven)$coefficients[,4][2]), digits = 4))); 

#otu_vegan <- plot_grid(otu_rich_vegan, otu_simpseven_vegan, otu_shannon_vegan, otu_invsimps_vegan,
#                       labels = c("A", "B", "C", "D"), 
#                       align = "h", nrow = 2, ncol = 2)
#otu_vegan
```


## Residual Analysis 
```{r check-lm-residuals, fig.width = 12, fig.height = 4, dependson = "plot-vegan-fracprod"}
##########################################################################
#############################   RESIDUALS   ##############################
##########################################################################

######################################################### RICHNESS
# Residual analysis of the RICHNESS Models
plot_residuals(lm_model = partprod_MLotu_rich, 
               lm_observed_y = filter(ML_otu_rich_stats, fraction == "WholePart")$frac_bacprod,
               main_title = "Particle-Associated Richness")


######################################################### SHANNON ENTROPY
# Residual analysis of the SHANNON ENTROPY Models
plot_residuals(lm_model = partprod_MLotu_shannon, 
               lm_observed_y = filter(ML_otu_shannon_stats, fraction == "WholePart")$frac_bacprod,
               main_title = "Particle-Associated Shannon")


######################################################### INVERSE SIMPSON
# Residual analysis of the INVERSE SIMPSON Models
plot_residuals(lm_model = partprod_MLotu_invsimps, 
               lm_observed_y = filter(ML_otu_invsimps_stats, fraction == "WholePart")$frac_bacprod,
               main_title = "Particle-Associated Inverse Simpson")


######################################################### SIMPSONS EVENNESS
# Residual analysis of the INVERSE SIMPSON Models
plot_residuals(lm_model = partprod_MLotu_simpseven, 
               lm_observed_y = filter(ML_otu_simpseven_stats, fraction == "WholePart")$frac_bacprod,
               main_title = "Particle-Associated Simpson's Evenness")
```


## Correlations 

```{r correlations, dependson = "plot-vegan-fracprod"}
##########################################################################
###########################   CORRELATIONS   #############################
##########################################################################
# RICHNESS vs SHANNON
cor(filter(ML_otu_rich_stats, fraction == "WholePart")$mean,
    filter(ML_otu_shannon_stats, fraction == "WholePart")$mean) # YES

# SHANNON VS INVERSE SIMPSON
cor(filter(ML_otu_shannon_stats, fraction == "WholePart")$mean,
    filter(ML_otu_invsimps_stats, fraction == "WholePart")$mean) # YES

# INVERSE SIMPSON VS SIMPSONS EVENNESS
cor(filter(ML_otu_invsimps_stats, fraction == "WholePart")$mean,
    filter(ML_otu_simpseven_stats, fraction == "WholePart")$mean) # YES

# SIMPSONS EVENNESS VS RICHNESS
cor(filter(ML_otu_simpseven_stats, fraction == "WholePart")$mean,
    filter(ML_otu_rich_stats, fraction == "WholePart")$mean) # YES

```


## Post-hoc analysis 
```{r post-hoc, dependson="vean-otu-alphadiv-calc"}
# Are the fractions different from each other in predicting fraction production?
prod_fracprod_values <- subset(otu_alphadiv, limnion == "Top" & year == "2015" & 
                          fraction == "WholePart" &
                          measure == "Richness") %>%
  dplyr::select(norep_filter_name, frac_bacprod) 

# Create a matrix with the 4 different diversity values 
prod_alpha <- subset(otu_alphadiv, limnion == "Top" & year == "2015" & 
                          fraction %in% c("WholePart", "WholeFree")) %>%
  dplyr::select(norep_filter_name, measure, mean) %>%
  spread(measure, mean)
row.names(prod_alpha) <- prod_alpha$norep_filter_name
prod_alpha$norep_filter_name = NULL
prod_alpha <- as.matrix(prod_alpha)

# Scale to a mean = 0 and  SD = 1
scale_prod_alphadiv <- scale(prod_alpha)

# Sanity Check
colMeans(scale_prod_alphadiv)  # faster version of apply(scaled.dat, 2, mean)
apply(scale_prod_alphadiv, 2, sd)

# Melt the data frame to be in long format
gather_prod_alpha <- as.data.frame(scale_prod_alphadiv) %>%   # Make scaled values a dataframe
  tibble::rownames_to_column(var = "norep_filter_name") %>%   # Add the rownames to keep samplenames
  gather(measure, mean, 2:5)                                  # Gather 4 columns and put values into 2
  
# Put it all together into one dataframe with 4 columns: sample_name, measure, mean, frac_bacprod 
prod_alpha_fracprod <- inner_join(gather_prod_alpha, prod_fracprod_values, by = "norep_filter_name") %>%
  mutate(measure = as.factor(measure))



# Double check values from above models
lm_by_divmeasure <- lm(frac_bacprod ~ mean/measure, data = prod_alpha_fracprod)
summary(lm_by_divmeasure)

# Run a post-hoc test
library(multcomp)
post_hoc_measure <- glht(lm_by_divmeasure, linfct = mcp(measure = "Tukey", interaction_average=TRUE),
                vcov=vcovHC(lm_by_divmeasure, type = "HC0"))
summary(post_hoc_measure)
detach("package:multcomp", unload=TRUE) # This package masks the dplyr select function = :(
```



# Per cell Fraction Production vs Diversity
```{r fracprod_percell-vs-div, eval = TRUE, fig.width = 15, fig.height = 4, warning = FALSE, dependson="subset-diversity"}
#########################################################  RICHNESS 
# Free-Living Richness vs fractional production per cell 
freeprod_percell_ML_otu_rich <- lm(log10(fracprod_per_cell) ~ mean, data = filter(ML_otu_rich_stats, fraction == "WholeFree"))
summary(freeprod_percell_ML_otu_rich)

# Particle-Associated Richness vs fractional production per cell 
partprod_percell_MLotu_rich <- lm(log10(fracprod_per_cell) ~ mean, data = filter(filter(ML_otu_rich_stats, fraction == "WholePart" & fracprod_per_cell != Inf)))
summary(partprod_percell_MLotu_rich)

# Plot 
rich_vs_fracprod_percell <- ggplot(filter(ML_otu_rich_stats, fracprod_per_cell != Inf),
       aes(x=mean, y=log10(fracprod_per_cell), color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_x_continuous(limits = c(180, 810), breaks = c(200, 400, 600, 800)) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Observed Richness") +
  geom_smooth(data=subset(ML_otu_rich_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 500, y=-8, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_percell_ML_otu_rich)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_percell_ML_otu_rich)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 650, y=-7, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_percell_MLotu_rich)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_percell_MLotu_rich)$coefficients[,4][2]), digits = 3)));


#########################################################  SHANNON
# Free-Living Shannon vs fractional production per cell 
freeprod_percell_ML_otu_shannon <- lm(log10(fracprod_per_cell) ~ mean, data = filter(ML_otu_shannon_stats, fraction == "WholeFree"))
summary(freeprod_percell_ML_otu_shannon)


# Particle-Associated Shannon vs fractional production per cell 
partprod_percell_MLotu_shannon <- lm(log10(fracprod_per_cell) ~ mean, 
                                     data = filter(filter(ML_otu_shannon_stats, fraction == "WholePart" & fracprod_per_cell != Inf)))
summary(partprod_percell_MLotu_shannon)

# Plot 
shannon_vs_fracprod_percell <- ggplot(filter(ML_otu_shannon_stats, fracprod_per_cell != Inf),
       aes(x=mean, y=log10(fracprod_per_cell), color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Shannon Entropy") +
  scale_x_continuous(limits = c(3.4, 5.85), breaks = c(3.5, 4, 4.5, 5, 5.5)) + 
  geom_smooth(data=subset(ML_otu_shannon_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 4.75, y=-8, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_percell_ML_otu_shannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_percell_ML_otu_shannon)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 5.5, y=-7, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_percell_MLotu_shannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_percell_MLotu_shannon)$coefficients[,4][2]), digits = 3)));


#########################################################  INVERSE SIMPSON 
# Free-Living Inverse Simpson vs fractional production per cell 
freeprod_percell_ML_otu_invsimps <- lm(log10(fracprod_per_cell) ~ mean, 
                                       data = filter(ML_otu_invsimps_stats, fraction == "WholeFree"))
summary(freeprod_percell_ML_otu_invsimps)

# Particle-Associated Inverse Simpson vs fractional production per cell 
partprod_percell_MLotu_invsimps <- lm(log10(fracprod_per_cell)  ~ mean, 
                              data = filter(ML_otu_invsimps_stats, fraction == "WholePart" & fracprod_per_cell != Inf))
summary(partprod_percell_MLotu_invsimps)

# Plot Simpson's Evenness
invsimps_vs_fracprod_percell <- ggplot(filter(ML_otu_invsimps_stats, fracprod_per_cell != Inf),
       aes(x=mean, y=log10(fracprod_per_cell) , color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  scale_x_continuous(limits = c(0,95), breaks = c(0, 20, 40, 60, 80), expand = c(0,0)) + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Inverse Simpson") +
  geom_smooth(data=subset(ML_otu_invsimps_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 50, y=-8, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_percell_ML_otu_invsimps)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_percell_ML_otu_invsimps)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 75, y=-7, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_percell_MLotu_invsimps)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_percell_MLotu_invsimps)$coefficients[,4][2]), digits = 3))); 



#########################################################  SIMPSON'S EVENNESS
# Free-Living Inverse Simpson vs fractional production per cell 
freeprod_percell_ML_otu_simpseven <- lm(log10(fracprod_per_cell) ~ mean, 
                                       data = filter(ML_otu_simpseven_stats, fraction == "WholeFree"))
summary(freeprod_percell_ML_otu_simpseven)

# Particle-Associated Inverse Simpson vs fractional production per cell 
partprod_percell_MLotu_simpseven <- lm(log10(fracprod_per_cell)  ~ mean, 
                              data = filter(ML_otu_simpseven_stats, fraction == "WholePart" & fracprod_per_cell != Inf))
summary(partprod_percell_MLotu_simpseven)

# Plot Simpson's Evenness
simpseven_vs_fracprod_percell <- ggplot(filter(ML_otu_simpseven_stats, fracprod_per_cell != Inf),
       aes(x=mean, y=log10(fracprod_per_cell) , color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), 
                     limits = c("WholePart", "WholeFree"),
                     breaks=c("WholePart", "WholeFree"),
                     labels=c("Particle-Associated", "Free-Living")) + 
  scale_x_continuous(limits = c(0.03,0.151), breaks = c(0.05, 0.1, 0.15))  +
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Simpson's Evenness") +
  geom_smooth(data=subset(ML_otu_simpseven_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 0.11, y=-8, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_percell_ML_otu_simpseven)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_percell_ML_otu_simpseven)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x =0.13, y=-7, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_percell_MLotu_simpseven)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_percell_MLotu_simpseven)$coefficients[,4][2]), digits = 3))); 



plot_grid(rich_vs_fracprod_percell + theme(legend.position= "none"), 
          shannon_vs_fracprod_percell + theme(legend.position= "none"),  
          invsimps_vs_fracprod_percell + theme(legend.position= "none"),
          simpseven_vs_fracprod_percell + theme(legend.position= c(0.35,0.9)),
          labels = c("A", "B", "C", "D"), 
          ncol = 4)
```



## Post-hoc analysis 
```{r percell-post-hoc, dependson=c("vean-otu-alphadiv-calc", "fracprod_percell-vs-div", "post-hoc")}
# Are the fractions different from each other in predicting fraction production?
prod_fracprodpercell_values <- subset(otu_alphadiv, limnion == "Top" & year == "2015" & 
                          fraction == "WholePart" &
                          measure == "Richness") %>%
  dplyr::select(norep_filter_name, fracprod_per_cell_noinf) 

# Melt the data frame to be in long format
gather_prod_alpha <- as.data.frame(scale_prod_alphadiv) %>%   # Make scaled values a dataframe
  tibble::rownames_to_column(var = "norep_filter_name") %>%   # Add the rownames to keep samplenames
  gather(measure, mean, 2:5)                                  # Gather 4 columns and put values into 2
  
# Put it all together into one dataframe with 4 columns: sample_name, measure, mean, frac_bacprod 
prod_alpha_fracprod_percell <- inner_join(gather_prod_alpha, prod_fracprodpercell_values, by = "norep_filter_name") %>%
  mutate(measure = as.factor(measure)) %>%
  dplyr::filter(!is.na(fracprod_per_cell_noinf))



# Double check values from above models
lm_percell_by_divmeasure <- lm(log10(fracprod_per_cell_noinf) ~ mean/measure, data = prod_alpha_fracprod_percell)
summary(lm_percell_by_divmeasure)

# Run a post-hoc test
library(multcomp)
post_hoc_measure <- glht(lm_percell_by_divmeasure, linfct = mcp(measure = "Tukey", interaction_average=TRUE),
                vcov=vcovHC(lm_percell_by_divmeasure, type = "HC0"))
summary(post_hoc_measure)
detach("package:multcomp", unload=TRUE) # This package masks the dplyr select function = :(
```


# Figure 2
```{r rich-and-invsimps-BEF, warning=FALSE, fig.width = 8, fig.height = 10, dependson= c("boxplot-diversity-comparison","plot-vegan-fracprod", "fracprod_percell-vs-div")}
plot_grid(poster_rich1 + xlab("\n Fraction \n") + ylab("Observed Richness") + 
                theme(legend.position = "none", axis.text.y = element_blank()) +coord_flip() +
                annotate("text", x=1.5, y=700, fontface = "bold",  size = 4, color = "gray40",
                          label= paste("***\np =", round(rich_wilcox$p.value, digits = 3))), 
          poster_invsimps1 + xlab("\n Fraction \n") + ylab("Inverse Simpson") + 
                theme(legend.position = c(0.78, 0.90), 
                      axis.text.y = element_blank(), 
                      legend.title = element_blank()) +
                coord_flip() +
                annotate("text", x=1.5, y=78, fontface = "bold",  size = 4, color = "gray40",
                          label= paste("NS\np =", round(simpson_wilcox$p.value, digits = 2))), 
          otu_rich_vegan +  ylab("Heterotrophic Production \n(μgC/L/hr)") + 
                theme(legend.position = "none"), 
          otu_invsimps_vegan + ylab("Heterotrophic Production \n(μgC/L/hr)") +
                theme(legend.position = c(0.7, 0.80)), 
          rich_vs_fracprod_percell + theme(legend.position= "none"), 
          invsimps_vs_fracprod_percell + theme(legend.position= "none"), 
          labels = c("A", "B", "C", "D", "E", "F"),
          ncol = 2, nrow = 3)
```


```{r shannon-and-simpseven-BEF, warning=FALSE, fig.width = 8, fig.height = 10, dependson= c("boxplot-diversity-comparison","plot-vegan-fracprod", "fracprod_percell-vs-div")}
plot_grid(poster_shannon1 + xlab("\n Fraction  \n") + ylab("Shannon Entropy") + 
                theme(legend.position = c(0.78, 0.90), axis.text.y = element_blank()) +
                coord_flip() +
                annotate("text", x=1.5, y=5.45, fontface = "bold",  size = 4, color = "gray40",
                          label= paste("**\np =", round(shannon_wilcox$p.value, digits = 3))),  
          poster_simpseven1 + xlab("\n Fraction \n") + ylab("Simpson's Evenness") + 
                theme(legend.position = "none", axis.text.y = element_blank()) +
                coord_flip() +
                annotate("text", x=1.5, y=0.135, fontface = "bold",  size = 4, color = "gray40",
                          label= paste("NS\np =", round(simpseven_wilcox$p.value, digits = 2))), 
          otu_shannon_vegan + ylab("Heterotrophic Production \n(μgC/L/hr)") + 
                theme(legend.position = c(0.7, 0.75)), 
          otu_simpseven_vegan + ylab("Heterotrophic Production \n(μgC/L/hr)") + 
                theme(legend.position = "none"), 
          shannon_vs_fracprod_percell + theme(legend.position= "none"), 
          simpseven_vs_fracprod_percell + theme(legend.position= "none"), 
          labels = c("A", "B", "C", "D", "E", "F"),
          ncol = 2, nrow = 3)
```



# Fraction Production Altogether
```{r fracprod-together-vs-div, eval = TRUE, fig.width = 15, fig.height = 8, warning = FALSE, dependson="subset-diversity"}
#########################################################  RICHNESS 
# Free-Living Richness vs fractional production per cell 
total_prodpercell_rich <- lm(log10(fracprod_per_cell) ~ mean, data = filter(filter(ML_otu_rich_stats, fracprod_per_cell != Inf)))
summary(total_prodpercell_rich)

anova(partprod_percell_MLotu_rich, total_prodpercell_rich)

# Plot 
combined_richness <- ggplot(filter(ML_otu_rich_stats, fracprod_per_cell != Inf), aes(x=mean, y=log10(fracprod_per_cell))) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Observed Richness") +
  geom_smooth(data= ML_otu_rich_stats, method='lm', color = "black") + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 500, y=-8, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(total_prodpercell_rich)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(total_prodpercell_rich)$coefficients[,4][2]), digits = 6)))



#########################################################  SHANNON
total_prodpercell_shannon <- lm(log10(fracprod_per_cell) ~ mean, data = filter(filter(ML_otu_shannon_stats,fracprod_per_cell != Inf)))
summary(total_prodpercell_shannon)

# Plot 
combined_shannon <- ggplot(filter(ML_otu_shannon_stats, fracprod_per_cell != Inf),
       aes(x=mean, y=log10(fracprod_per_cell))) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Shannon Entropy") +
  geom_smooth(data= ML_otu_shannon_stats, method='lm', color = "black") + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 4.75, y=-8, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(total_prodpercell_shannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(total_prodpercell_shannon)$coefficients[,4][2]), digits = 6)));


#########################################################  INVERSE SIMPSON 
total_prodpercell_simpson <- lm(log10(fracprod_per_cell)  ~ mean, data = filter(ML_otu_invsimps_stats, fracprod_per_cell != Inf))
summary(total_prodpercell_simpson)

# Plot Simpson's Evenness
combined_invsimps <- ggplot(filter(ML_otu_invsimps_stats, fracprod_per_cell != Inf),
       aes(x=mean, y=log10(fracprod_per_cell))) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  scale_x_continuous(limits = c(0,100), expand = c(0,0)) + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Inverse Simpson") +
  geom_smooth(data = ML_otu_invsimps_stats, method='lm', color = "black") + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 50, y=-8, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(total_prodpercell_simpson)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(total_prodpercell_simpson)$coefficients[,4][2]), digits = 5))) 


#########################################################  SIMPSON'S EVENNESS
total_prodpercell_simpseven <- lm(log10(fracprod_per_cell) ~ mean, 
                                       data = filter(ML_otu_simpseven_stats, fracprod_per_cell != Inf))
summary(total_prodpercell_simpseven)


# Plot Simpson's Evenness
combined_simpseven <- ggplot(filter(ML_otu_simpseven_stats, fracprod_per_cell != Inf),
       aes(x=mean, y=log10(fracprod_per_cell))) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_x_continuous(limits = c(0.03,0.151), breaks = c(0.05, 0.1, 0.15))  +
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Simpson's Evenness") +
  geom_smooth(data = ML_otu_simpseven_stats, method='lm', color = "black") + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 0.11, y=-8, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(total_prodpercell_simpseven)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(total_prodpercell_simpseven)$coefficients[,4][2]), digits = 2)))



plot_grid(rich_vs_fracprod_percell + ylab("log10(Production/Cell)\n (μgC/cell/hr)") + theme(legend.position = "none"), 
          shannon_vs_fracprod_percell + ylab("log10(Production/Cell)\n (μgC/cell/hr)")+ theme(legend.position = "none"), 
          invsimps_vs_fracprod_percell + ylab("log10(Production/Cell)\n (μgC/cell/hr)")+ theme(legend.position = "none"), 
          simpseven_vs_fracprod_percell + ylab("log10(Production/Cell)\n (μgC/cell/hr)") +
                    theme(legend.position = c(0.35,0.9)), 
          combined_richness, combined_shannon, combined_invsimps,combined_simpseven, 
          labels = c("A", "B", "C", "D","E", "F", "G", "H"), 
          ncol = 4, nrow = 2)
```



```{r plot-combined-percell, eval = TRUE, fig.width = 15, fig.height = 4, warning = FALSE, dependson=c("subset-diversity", "fracprod-together-vs-div")}

plot_grid(combined_richness, combined_shannon, combined_invsimps,combined_simpseven, 
          labels = c("A", "B", "C", "D"), 
          ncol = 4, nrow = 1)
```


# Exponential Shannon
```{r exp-shannon, eval = TRUE, fig.width = 10, fig.height = 4, warning = FALSE, dependson="subset-diversity"}
#########################################################  EXPONENTIALSHANNON
# Free-Living EXPONENTIAL Shannon vs fractional production per cell 
freeprod_expshannon <- lm(frac_bacprod ~ exp(mean), data = filter(ML_otu_shannon_stats, fraction == "WholeFree"))
summary(freeprod_expshannon)


# Particle-Associated EXPONENTIAL Shannon vs fractional production per cell 
partprod_expshannon <- lm(frac_bacprod ~ exp(mean),
                                        data = filter(filter(ML_otu_shannon_stats, fraction == "WholePart" & fracprod_per_cell != Inf)))
summary(partprod_expshannon)

# Plot 
expshannon_fracprod <- ggplot(filter(ML_otu_shannon_stats, fracprod_per_cell != Inf),
       aes(x= exp(mean), y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + #geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  ylab("Fraction Production (μgC/L/hr)") + 
  xlab("Exponential Shannon") +
  #scale_x_continuous(limits = c(3.4, 5.85), breaks = c(3.5, 4, 4.5, 5, 5.5)) + 
  geom_smooth(data=subset(ML_otu_shannon_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 150, y=35, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_expshannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_expshannon)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 200, y=7, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_expshannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_expshannon)$coefficients[,4][2]), digits = 3)));




#########################################################  EXPONENTIALSHANNON
# Free-Living EXPONENTIAL Shannon vs fractional production per cell 
freeprod_percell_ML_otu_expshannon <- lm(log10(fracprod_per_cell) ~ exp(mean), data = filter(ML_otu_shannon_stats, fraction == "WholeFree"))
summary(freeprod_percell_ML_otu_expshannon)


# Particle-Associated EXPONENTIAL Shannon vs fractional production per cell 
partprod_percell_MLotu_expshannon <- lm(log10(fracprod_per_cell) ~ exp(mean), 
                                     data = filter(filter(ML_otu_shannon_stats, fraction == "WholePart" & fracprod_per_cell != Inf)))
summary(partprod_percell_MLotu_expshannon)

# Plot 
expshannon_fracprod_percell <- ggplot(filter(ML_otu_shannon_stats, fracprod_per_cell != Inf),
       aes(x= exp(mean), y=log10(fracprod_per_cell), color = fraction)) + 
  geom_point(size = 3.5) + #geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  ylab("log10(Fraction Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Exponential Shannon") +
  #scale_x_continuous(limits = c(3.4, 5.85), breaks = c(3.5, 4, 4.5, 5, 5.5)) + 
  geom_smooth(data=subset(ML_otu_shannon_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 150, y=-8, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_percell_ML_otu_expshannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(freeprod_percell_ML_otu_expshannon)$coefficients[,4][2]), digits = 2))) + 
  annotate("text", x = 200, y=-7, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_percell_MLotu_expshannon)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(partprod_percell_MLotu_expshannon)$coefficients[,4][2]), digits = 3)));


#########################################################  EXPONENTIALSHANNON
combined_expshannon_percell <- lm(log10(fracprod_per_cell) ~ exp(mean), data = filter(filter(ML_otu_shannon_stats, fracprod_per_cell != Inf)))
summary(combined_expshannon_percell)

# Plot 
expshannon_combined <- ggplot(filter(ML_otu_shannon_stats, fracprod_per_cell != Inf),
       aes(x= exp(mean), y=log10(fracprod_per_cell))) + 
  geom_point(size = 3.5) + 
  ylab("log10(Fraction Production/Cell)\n (μgC/cell/hr)") + 
  xlab("Exponential Shannon") +
  #scale_x_continuous(limits = c(3.4, 5.85), breaks = c(3.5, 4, 4.5, 5, 5.5)) + 
  geom_smooth(data= ML_otu_shannon_stats, method='lm', color = "black") + 
  theme(legend.position=c(0.2,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 150, y=-8, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(combined_expshannon_percell)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(combined_expshannon_percell)$coefficients[,4][2]), digits = 5))) 

plot_grid(expshannon_fracprod, expshannon_fracprod_percell, expshannon_combined,
          labels = c("A", "B", "C"), ncol = 3)


```





# Phylogenetic Community Structure 

# Faiths PD
```{r faiths-pd, fig.width=4, fig.height=4}
# Create the metadata frame 
nosed_meta_data <- nosed_meta_data %>%
  mutate(fraction_bac_abund = as.numeric(fraction_bac_abund),
         fracprod_per_cell = frac_bacprod/(1000*fraction_bac_abund),
         fracprod_per_cell_noinf = ifelse(fracprod_per_cell == Inf, NA, fracprod_per_cell))



## Calculate Faith's PD and species richness for sample 
#faiths_pd_RAREFIED <- pd(relabund_otu_RAREFIED, tree_RAREFIED_rm10, include.root = FALSE)
#faiths_pd_RAREFIED$norep_filter_name <- row.names(faiths_pd_RAREFIED)
  
#write.csv(faiths_pd_RAREFIED, file = "PrunedTree/mpd_mntd/faithsPD_rarefy_rm10.csv", row.names = FALSE)
faiths_pd_RAREFIED_rm10 <- read.csv("PrunedTree/mpd_mntd/faithsPD_rarefy_rm10.csv", header = TRUE)

# Join Faith's PD with the rest of the metadata 
meta_data_PD <- left_join(faiths_pd_RAREFIED_rm10, nosed_meta_data, by = "norep_filter_name")


### Is there a correlation between species richness and faith's PD?
lm_PD_vs_SR <- lm(PD ~ SR, data = meta_data_PD)
ggplot(meta_data_PD, aes(y = PD, x = SR)) + 
  geom_point(size = 3) + ylab("Faith's Phylogenetic Diversity") + 
  xlab("Species richness") +
  geom_smooth(method = "lm") +
    annotate("text", x = 300, y=15, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_PD_vs_SR)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(lm_PD_vs_SR)$coefficients[,4][2]), digits = 35))) 
```


##Taxa.Labels
```{r taxa-labels,  fig.width= 14, fig.height = 7, dependson = "faiths-pd"}
### UNWEIGHTED
unweighted_sesMPD_taxalab <- read.csv("PrunedTree/mpd_mntd/unweighted_sesMPD_taxalab.csv", header = TRUE) %>%
  dplyr::rename(norep_filter_name = X) %>%
  left_join(nosed_meta_data, by = "norep_filter_name") %>%
  # Create discrete pvalues and reorder factors for fraction and lakesite
  mutate(pval = ifelse(mpd.obs.p > 0.9499, "high_pval",
                       ifelse(mpd.obs.p < 0.0511, "low_pval",
                              "insignificant")),
         fraction = factor(fraction, levels = c("WholePart", "Particle", "WholeFree", "Free")),
         lakesite = factor(lakesite, levels =c("MOT", "MDP", "MBR", "MIN")))


### WEIGHTED
WEIGHTED_sesMPD_taxalab <- read.csv("PrunedTree/mpd_mntd/weighted_sesMPD_taxalab.csv", header = TRUE) %>%
  dplyr::rename(norep_filter_name = X) %>%
  left_join(nosed_meta_data, by = "norep_filter_name") %>%
  # Create discrete pvalues and reorder factors for fraction and lakesite
  mutate(pval = ifelse(mpd.obs.p > 0.9499, "high_pval",
                       ifelse(mpd.obs.p < 0.0511, "low_pval",
                              "insignificant")),
         fraction = factor(fraction, levels = c("WholePart", "Particle", "WholeFree", "Free")),
         lakesite = factor(lakesite, levels =c("MOT", "MDP", "MBR", "MIN")))

p1 <- ggplot(unweighted_sesMPD_taxalab, 
             aes(x = lakesite, y = mpd.obs.z, color = pval, fill = fraction)) + 
  geom_point(size = 3, position = position_jitterdodge()) + 
  ggtitle("Unweighted MPD") +
  ylab("Mean Pairwise Distance \n (ses.mpd)") +
  geom_boxplot(alpha = 0.5, color = "black") +
  scale_color_manual(values = pd_colors) + 
  scale_fill_manual(values = fraction_colors) +
  facet_grid(.~fraction, scale = "free_x") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))

p2 <- ggplot(WEIGHTED_sesMPD_taxalab, 
             aes(x = lakesite, y = mpd.obs.z, color = pval, fill = fraction)) + 
  geom_point(size = 3, position = position_jitterdodge()) + 
  ggtitle("Weighted MPD") +
  ylab("Mean Pairwise Distance \n (ses.mpd)") +
  geom_boxplot(alpha = 0.5, color = "black") +
  scale_color_manual(values = pd_colors) + 
  scale_fill_manual(values = fraction_colors) +
  facet_grid(.~fraction, scale = "free_x") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))


### UNWEIGHTED
unweighted_sesMNTD_taxalab <- read.csv("PrunedTree/mpd_mntd/unweighted_sesMNTD_taxalab.csv", header = TRUE) %>%
  dplyr::rename(norep_filter_name = X) %>%
  left_join(nosed_meta_data, by = "norep_filter_name") %>%
  # Create discrete pvalues and reorder factors for fraction and lakesite
  mutate(pval = ifelse(mntd.obs.p > 0.9499, "high_pval",
                       ifelse(mntd.obs.p < 0.0511, "low_pval",
                              "insignificant")),
         fraction = factor(fraction, levels = c("WholePart", "Particle", "WholeFree", "Free")),
         lakesite = factor(lakesite, levels =c("MOT", "MDP", "MBR", "MIN")))


### WEIGHTED
WEIGHTED_sesMNTD_taxalab <- read.csv("PrunedTree/mpd_mntd/weighted_sesMNTD_taxalab.csv", header = TRUE) %>%
  dplyr::rename(norep_filter_name = X) %>%
  left_join(nosed_meta_data, by = "norep_filter_name") %>%
  # Create discrete pvalues and reorder factors for fraction and lakesite
  mutate(pval = ifelse(mntd.obs.p > 0.9499, "high_pval",
                       ifelse(mntd.obs.p < 0.0511, "low_pval",
                              "insignificant")),
         fraction = factor(fraction, levels = c("WholePart", "Particle", "WholeFree", "Free")),
         lakesite = factor(lakesite, levels =c("MOT", "MDP", "MBR", "MIN")))

p3 <- ggplot(unweighted_sesMNTD_taxalab, 
             aes(x = lakesite, y = mntd.obs.z, color = pval, fill = fraction)) + 
  geom_point(size = 3, position = position_jitterdodge()) + 
  ggtitle("Unweighted MNTD") +
  ylab("Mean Nearest Taxon Distance \n (ses.mntd)") +
  geom_boxplot(alpha = 0.5, color = "black") +
  scale_color_manual(values = pd_colors) + 
  scale_fill_manual(values = fraction_colors) +
  facet_grid(.~fraction, scale = "free_x") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))

p4 <- ggplot(WEIGHTED_sesMNTD_taxalab, 
             aes(x = lakesite, y = mntd.obs.z, color = pval, fill = fraction)) + 
  geom_point(size = 3, position = position_jitterdodge()) + 
  ggtitle("Weighted MNTD") +
  ylab("Mean Nearest Taxon Distance \n (ses.mntd)") +
  geom_boxplot(alpha = 0.5, color = "black") +
  scale_color_manual(values = pd_colors) + 
  scale_fill_manual(values = fraction_colors) +
  facet_grid(.~fraction, scale = "free_x") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))

# Put all of it together into one plot
plot_grid(p1, p2, p3, p4,
          labels = c("A", "B", "C", "D"), 
          ncol = 2, nrow = 2)
```

```{r taxalab-vs-fracprod, fig.width= 10, fig.height = 10}
##########  MPD ANALYSIS 
lmFL_unweightedMPD_taxalab <- lm(frac_bacprod ~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction == "WholeFree"))
lmPA_unweightedMPD_taxalab <- lm(frac_bacprod ~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction == "WholePart"))

unweightedMPD_vs_fracprod_taxlab <- ggplot(filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y = frac_bacprod, color = fraction)) + 
  geom_point(size = 3) +
  xlab("Unweighted Mean Pairwise Distance \n (ses.mpd)") + 
  ylab("Heterotrophic Production \n(μgC/L/hr)") +
  ggtitle("Unweighted MPD: Taxa Labels") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm") +
  theme(legend.position = c(0.2, 0.87), legend.title = element_blank()) +
    annotate("text", x = -0, y=-2, 
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(lmPA_unweightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lmPA_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) +
  annotate("text", x = 2, y=60, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(lmFL_unweightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
                       "p =", round(unname(summary(lmFL_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) 


####### WEIGHTED MPD ANALYSIS
lmFL_weightedMPD_taxalab <- lm(frac_bacprod ~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction == "WholeFree"))
lmPA_weightedMPD_taxalab <- lm(frac_bacprod ~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction == "WholePart"))

weightedMPD_vs_fracprod_taxlab <- ggplot(filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y = frac_bacprod, color = fraction)) + 
  geom_point(size = 3) +
  xlab("Weighted Mean Pairwise Distance  \n (ses.mpd)") + 
  ylab("Heterotrophic Production \n(μgC/L/hr)") +
  ggtitle("Weighted MPD: Taxa Labels") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  #geom_smooth(method = "lm", data = filter(WEIGHTED_sesMPD_taxalab, fraction == "WholePart")) +
  theme(legend.position = c(0.2, 0.87), legend.title = element_blank()) +
    annotate("text", x = 1, y=25, 
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(lmPA_weightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lmPA_weightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) +
  annotate("text", x = 0, y=45, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(lmFL_weightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
                       "p =", round(unname(summary(lmFL_weightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) 


####### UNWEIGHTED MNTD ANALYSIS
lmFL_unweightedMNTD_taxalab <- lm(frac_bacprod ~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction == "WholeFree"))
lmPA_unweightedMNTD_taxalab <- lm(frac_bacprod ~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction == "WholePart"))

unweightedMNTD_vs_fracprod_taxlab <- ggplot(filter(unweighted_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mntd.obs.z, y = frac_bacprod, color = fraction)) + 
  geom_point(size = 3) +
  xlab("Unweighted Mean Nearest Taxon Distance  \n (ses.mntd)") + 
  ylab("Heterotrophic Production \n(μgC/L/hr)") +
  ggtitle("Unweighted MNTD: Taxa Labels") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm", data = filter(unweighted_sesMNTD_taxalab, fraction == "WholePart")) +
  theme(legend.position = c(0.87, 0.87), legend.title = element_blank()) +
    annotate("text", x = -3.6, y=-2, 
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(lmPA_unweightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lmPA_unweightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) +
  annotate("text", x = -1.7, y=60, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(lmFL_unweightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
                       "p =", round(unname(summary(lmFL_unweightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) 


####### WEIGHTED MPD ANALYSIS
lmFL_weightedMNTD_taxalab <- lm(frac_bacprod ~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholeFree"))
lmPA_weightedMNTD_taxalab <- lm(frac_bacprod ~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholePart"))

weightedMNTD_vs_fracprod_taxlab <- ggplot(filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mntd.obs.z, y = frac_bacprod, color = fraction)) + 
  geom_point(size = 3) +
  xlab("Weighted Mean Nearest Taxon Distance  \n (ses.mntd)") + 
  ylab("Heterotrophic Production \n(μgC/L/hr)") +
  ggtitle("Weighted MNTD: Taxa Labels") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  #geom_smooth(method = "lm", data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholePart")) +
  theme(legend.position = c(0.87, 0.87), legend.title = element_blank()) +
    annotate("text", x = -1.25, y=25, 
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(lmPA_weightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lmPA_weightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) +
  annotate("text", x = -1.25, y=45, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(lmFL_weightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
                       "p =", round(unname(summary(lmFL_weightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) 

# Plot it altogether 
plot_grid(unweightedMPD_vs_fracprod_taxlab, unweightedMNTD_vs_fracprod_taxlab, 
          weightedMPD_vs_fracprod_taxlab, weightedMNTD_vs_fracprod_taxlab,
          labels = c("A", "B", "C", "D"), 
          ncol = 2, nrow = 2)
```


```{r taxalab-comparison, fig.width= 12, fig.height = 5}
unweight_MPD_wilcox <- wilcox.test(mpd.obs.z ~ fraction, 
             data = filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")))
unweight_MPD_wilcox

# Make a data frame to draw significance line in boxplot (visually calculated)
mpd_nums1 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(3.2,3.3,3.3,3.2)) # WholePart vs WholeFree

taxlab_unweight_mpd <- ggplot(filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(y = mpd.obs.z, x = fraction)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 1.5) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  scale_x_discrete(breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free\nLiving", "Particle\nAssociated")) + 
  geom_point(size = 3, position = position_jitterdodge(), aes(color = fraction, fill = fraction)) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  ylab("Unweighted Mean Pairwise Dist") +
  xlab("") +
  geom_path(data = mpd_nums1, aes(x = a, y = b), linetype = 1, color = "gray40") +
  scale_y_continuous(limits = c(-1.5, 3.5), breaks = c(-1, 0, 1, 2, 3, 4)) +
  theme(legend.position = "none")



weight_MPD_wilcox <- wilcox.test(mpd.obs.z ~ fraction, 
             data = filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")))
weight_MPD_wilcox

filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")) %>%
  group_by(fraction) %>%
  summarize(mean(mpd.obs.z),  median(mpd.obs.z), sd(mpd.obs.z))

# Make a data frame to draw significance line in boxplot (visually calculated)
mpd_nums2 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(2.2,2.3,2.3,2.2)) # WholePart vs WholeFree

taxlab_weight_mpd <- ggplot(filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(y = mpd.obs.z, x = fraction)) +
  geom_hline(yintercept = 0, slope = 0, linetype = "dashed", size = 1.5) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  scale_x_discrete(breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free\nLiving", "Particle\nAssociated")) + 
  geom_point(size = 3, position = position_jitterdodge(), aes(color = fraction, fill = fraction)) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  xlab("") + 
  ylab("Weighted Mean Pairwise Dist") +
  geom_path(data = mpd_nums2, aes(x = a, y = b), linetype = 1, color = "gray40") +
  scale_y_continuous(limits = c(-3.5, 2.5), breaks = c(-3, -2, -1, 0, 1, 2, 3, 4)) +
  theme(legend.position = "none")


########################################## MNTD
###### UNWEIGHTED MNTD 
unweight_MNTD_wilcox <- wilcox.test(mntd.obs.z ~ fraction, 
             data = filter(unweighted_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")))
unweight_MNTD_wilcox

# Make a data frame to draw significance line in boxplot (visually calculated)
mntd_nums1 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(-0.5,-0.4,-0.4,-0.5)) # WholePart vs WholeFree

taxlab_unweight_mntd <- ggplot(filter(unweighted_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(y = mntd.obs.z, x = fraction)) +  
  geom_hline(yintercept = 0, slope = 0, linetype = "dashed", size = 1.5) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  scale_x_discrete(breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free\nLiving", "Particle\nAssociated")) + 
  geom_point(size = 3, position = position_jitterdodge(), aes(color = fraction, fill = fraction)) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  xlab("") + 
  ylab("Unweighted Mean Nearest Taxon Dist") +
  geom_path(data = mntd_nums1, aes(x = a, y = b), linetype = 1, color = "gray40") +
  scale_y_continuous(limits = c(-4, 0), breaks = c(-4, -3, -2, -1, 0)) +
  theme(legend.position = "none")




##### WEIGHTED MNTD
weight_MNTD_wilcox <- wilcox.test(mntd.obs.z ~ fraction, 
             data = filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")))
weight_MNTD_wilcox

filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")) %>%
  group_by(fraction) %>%
  summarize(mean(mntd.obs.z),  median(mntd.obs.z))

# Make a data frame to draw significance line in boxplot (visually calculated)
mntd_nums2 <- data.frame(a = c(1.15,1.15,1.85,1.85), b = c(1,1.1,1.1,1)) # WholePart vs WholeFree

taxlab_weight_mntd <- ggplot(filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(y = mntd.obs.z, x = fraction)) +
  geom_hline(yintercept = 0, slope = 0, linetype = "dashed", size = 1.5) +
  scale_fill_manual(values = fraction_colors, 
                    breaks=c("WholeFree", "WholePart"), 
                    labels=c("Free-Living", "Particle-Associated")) +
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free-Living", "Particle-Associated")) + 
  scale_x_discrete(breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free\nLiving", "Particle\nAssociated")) + 
  geom_point(size = 3, position = position_jitterdodge(), aes(color = fraction, fill = fraction)) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA, aes(color = fraction, fill = fraction)) +
  xlab("") + 
  ylab("Weighted Mean Nearest Taxon Dist") +
  geom_path(data = mntd_nums2, aes(x = a, y = b), linetype = 1, color = "gray40") +
  scale_y_continuous(limits = c(-3.25, 1.25), breaks = c(-3, -2, -1, 0, 1)) +
  theme(legend.position = "none")



plot_grid(taxlab_unweight_mpd + ylim(-4.5, 3.5) +
            annotate("text", x=1.55, y=max(mpd_nums1$b), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS\np =", round(unweight_MPD_wilcox$p.value, digits = 2))), 
          taxlab_unweight_mntd + ylim(-4.5, 3.5) +
            annotate("text", x=1.55, y=max(mntd_nums1$b), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS\np =", round(unweight_MNTD_wilcox$p.value, digits = 2))), 
          taxlab_weight_mpd + ylim(-4.5, 3.5) +
            annotate("text", x=1.55, y=max(mpd_nums2$b), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MPD_wilcox$p.value, digits = 6))),
          taxlab_weight_mntd + ylim(-4.5, 3.5) +
            annotate("text", x=1.55, y=max(mntd_nums2$b), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MNTD_wilcox$p.value, digits = 3))),
          labels = c("A", "B", "C", "D"),
          ncol = 4)
```



# Bulk Productivity
  
  
- Linear Models with **BULK** Productivity:  
      - Both PA and FL together:  
              - Unweighted MPD:  **NS**  
              - Weighted MPD: **NS** (R2 = 0.1061, p = 0.066)  
              - Unweighted MNTD:  **NS** (p = 0.089)  
              - Weighted MNTD: **Significant** (p = 0.01129, R2 = 0.2242)

```{r taxalab-production-BEF, fig.width= 15, fig.height = 4, dependson = "taxa-labels"}
##########  MPD ANALYSIS 
# Free-Living Model
prod_lmFL_unweightedMPD_taxalab <- lm(frac_bacprod~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction == "WholeFree"))
summary(prod_lmFL_unweightedMPD_taxalab)

# Particle-associated model
prod_lmPA_unweightedMPD_taxalab <- lm(frac_bacprod~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction == "WholePart"))
summary(prod_lmPA_unweightedMPD_taxalab)

# BOTH FRACRTIONS TOGETHER
prod_lmTOGET_unweightedMPD_taxalab <- lm(frac_bacprod~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(prod_lmTOGET_unweightedMPD_taxalab)


lm_3 <- lm(frac_bacprod~ mpd.obs.z/fraction, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(lm_3)

lm_3_minus1 <- lm(frac_bacprod~ mpd.obs.z/fraction - 1, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(lm_3_minus1)


plot_unweightedMPD_prod <-   ggplot(filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y =frac_bacprod, color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Unweighted Mean Pairwise Dist") + 
  ylab("Heterotrophic Production \n (μgC/L/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm") +
  theme(legend.position = "none", legend.title = element_blank()) +
  annotate("text", x = -3, y=20, 
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(prod_lmPA_unweightedMPD_taxalab)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(prod_lmPA_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 3))) +
  annotate("text", x = -3, y=60, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(prod_lmFL_unweightedMPD_taxalab)$adj.r.squared, digits = 2), "\n", 
                       "p =", round(unname(summary(prod_lmFL_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 3))) + 
  annotate("text", x = -3, y=-1, color = "black", fontface = "bold", label = paste("Combined = NS"))


ggplot(filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y =frac_bacprod, color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Unweighted Mean Pairwise Dist") + 
  ylab("Heterotrophic Production \n (μgC/L/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm", formula = y ~ x -1) +
  theme(legend.position = "none", legend.title = element_blank()) #+
  #annotate("text", x = -3, y=20, 
  #         color = "firebrick3", fontface = "bold",
  #         label = paste("R2 =", round(summary(prod_lmPA_unweightedMPD_taxalab)$adj.r.squared, digits = 2), "\n", 
  #                       "p =", round(unname(summary(prod_lmPA_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 3))) +
  #annotate("text", x = -3, y=60, 
  #       color = "cornflowerblue", fontface = "bold",
  #       label = paste("R2 =", round(summary(prod_lmFL_unweightedMPD_taxalab)$adj.r.squared, digits = 2), "\n", 
  #                     "p =", round(unname(summary(prod_lmFL_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 3))) + 
  #annotate("text", x = -3, y=-1, color = "black", fontface = "bold", label = paste("Combined = NS"))


############ WEIGHTED MPD
##########  MPD ANALYSIS 
# Free-Living Model
prod_lmFL_weightedMPD_taxalab <- lm(frac_bacprod~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction == "WholeFree"))
summary(prod_lmFL_weightedMPD_taxalab)

# Particle-associated model
prod_lmPA_weightedMPD_taxalab <- lm(frac_bacprod~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction == "WholePart"))
summary(prod_lmPA_weightedMPD_taxalab)


###  COMBINE BOTH PARTICLE AND FREE INTO ONE MODEL
prod_lmTOGET_weightedMPD_taxalab <- lm(frac_bacprod~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(prod_lmTOGET_weightedMPD_taxalab)

# Residual analysis of the UNWEIGHTED MPD Models: PARTICLE-ASSOCIATED
ALL_weight_sesMPD_df <- dplyr::filter(WEIGHTED_sesMPD_taxalab, 
                                      fraction %in% c("WholePart", "WholeFree") & !is.na(fracprod_per_cell_noinf))

plot_residuals(lm_model = prod_lmALL_weightedMPD_taxalab, 
               lm_observed_y = (ALL_weight_sesMPD_df$fracprod_per_cell_noinf),
               main_title = "Both Fractions Weighted MPD")



plot_weightedMPD_prod_TOGETHER <- ggplot(filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y =frac_bacprod)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3,  aes(color = fraction)) +
  xlab("Weighted Mean Pairwise Dist") + 
  ylab("Heterotrophic Production \n (μgC/L/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm", color = "black", linetype = "dashed") +
  theme(legend.position = "none") +
  annotate("text", x = -3, y=54, 
           color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(prod_lmTOGET_weightedMPD_taxalab)$adj.r.squared, digits = 3), "\n", 
                         "p =", round(unname(summary(prod_lmTOGET_weightedMPD_taxalab)$coefficients[,4][2]), digits = 3))) +
  annotate("text", x = -3, y=-1, color = "#CCBB51", fontface = "bold", label = paste("Separate = NS"))


plot_weightedMPD_prod <- ggplot(filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y =frac_bacprod, color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Weighted Mean Pairwise Dist") + 
  ylab("Heterotrophic Production \n (μgC/L/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  #geom_smooth(method = "lm") +
  theme(legend.position = "none", legend.title = element_blank()) #+
  #annotate("text", x = -2, y=-6.5, 
  #         color = "firebrick3", fontface = "bold",
  #         label = paste("R2 =", round(summary(prod_lmPA_weightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                       "p =", round(unname(summary(prod_lmPA_weightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) +
  #annotate("text", x = 0, y=-8.2, 
  #       color = "cornflowerblue", fontface = "bold",
  #       label = paste("R2 =", round(summary(prod_lmFL_weightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                     "p =", round(unname(summary(prod_lmFL_weightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) 





####### UNWEIGHTED MNTD ANALYSIS
# Free-living
prod_lmFL_unweightedMNTD_taxalab <- lm(frac_bacprod~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction == "WholeFree"))
summary(prod_lmFL_unweightedMNTD_taxalab)

# Particle-associated
prod_lmPA_unweightedMNTD_taxalab <- lm(frac_bacprod~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction == "WholePart"))
summary(prod_lmPA_unweightedMNTD_taxalab)

# TOGETHER
prod_lmTOGET_unweightedMNTD_taxalab <- lm(frac_bacprod~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(prod_lmTOGET_unweightedMNTD_taxalab)



plot_unweightedMNTD_prod <- ggplot(filter(unweighted_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mntd.obs.z, y = frac_bacprod, color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Unweighted Mean Nearest Taxon Dist") + 
  ylab("Heterotrophic Production \n (μgC/L/hr)") + 
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free", "Particle")) + 
  geom_smooth(method = "lm", data = filter(unweighted_sesMNTD_taxalab, fraction == "WholePart")) +
  theme(legend.position = c(0.87, 0.87), 
        legend.title = element_blank()) +
  annotate("text", x = 1.5, y=10,
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(prod_lmPA_unweightedMNTD_taxalab)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(prod_lmPA_unweightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) #+
  #annotate("text", x = -2.4, y=-7.8, 
  #       color = "cornflowerblue", fontface = "bold",
  #       label = paste("R2 =", round(summary(prod_lmFL_unweightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                     "p =", round(unname(summary(prod_lmFL_unweightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) 


####### WEIGHTED MPD ANALYSIS
# Free-living
prod_lmFL_weightedMNTD_taxalab <- lm(frac_bacprod~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholeFree"))
summary(prod_lmFL_weightedMNTD_taxalab)
# Particle-associated
prod_lmPA_weightedMNTD_taxalab <- lm(frac_bacprod~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholePart"))
summary(prod_lmPA_weightedMNTD_taxalab)

###  COMBINE BOTH PARTICLE AND FREE INTO ONE MODEL
prod_lmTOGET_weightedMNTD_taxalab <- lm(frac_bacprod~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(prod_lmTOGET_weightedMNTD_taxalab)  # p = 0.01129, R2 = 0.2242

# Residual analysis of the UNWEIGHTED MPD Models: PARTICLE-ASSOCIATED
ALL_weight_sesMNTD_df <- dplyr::filter(WEIGHTED_sesMNTD_taxalab, 
                                      fraction %in% c("WholePart", "WholeFree") & !is.na(fracprod_per_cell_noinf))


plot_weightedMNTD_prod <- ggplot(filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mntd.obs.z, y = frac_bacprod)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3, aes( color = fraction)) +
  xlab("Weighted Mean Nearest Taxon Dist") + 
  ylab("Heterotrophic Production \n (μgC/L/hr)") + 
  geom_smooth(method = "lm") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  #geom_smooth(method = "lm", data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholeFree")) +
  theme(legend.position = "none", legend.title = element_blank()) 
  #annotate("text", x = -2.5, y=-5.9,
  #         color = "firebrick3", fontface = "bold",
  #         label = paste("R2 =", round(summary(prod_lmPA_weightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                       "p =", round(unname(summary(prod_lmPA_weightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) +
  #annotate("text", x = 1.5, y=50, 
  #       color = "cornflowerblue", fontface = "bold",
  #       label = paste("R2 =", round(summary(prod_lmFL_weightedMNTD_taxalab)$adj.r.squared, digits = 2), "\n", 
  #                     "p =", round(unname(summary(prod_lmFL_weightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) 






plot_grid(plot_unweightedMPD_prod, plot_unweightedMNTD_prod,
          plot_weightedMPD_prod, plot_weightedMNTD_prod,
  labels = c("A", "B", "C", "D"),
  nrow= 1, ncol = 4
)

```


### Per-Cell Productivity

- Linear Models with **per cell** Productivity:  
      - Both PA and FL together:  
              - Unweighted MPD:  **Significant** (R2 = 0.5245, p = 5.631e-05)  
              - Weighted MPD: **Significant** (R2 = 0.2976, p = 0.004181)  
              - Unweighted MNTD:  **NS**  
              - Weighted MNTD: **NS**

```{r taxalab-percell, fig.width= 15, fig.height = 4, dependson = "taxa-labels"}
##########  MPD ANALYSIS 
# Free-Living Model
percell_lmFL_unweightedMPD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction == "WholeFree"))
summary(percell_lmFL_unweightedMPD_taxalab)

# Particle-associated model
percell_lmPA_unweightedMPD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction == "WholePart"))
summary(percell_lmPA_unweightedMPD_taxalab)

anova(percell_lmFL_unweightedMPD_taxalab, percell_lmPA_unweightedMPD_taxalab)



# BOTH FRACTIONS TOGETHER 
percell_lmTOGET_unweightedMPD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
                                      data = filter(unweighted_sesMPD_taxalab, fraction %in% c("WholeFree","WholePart")))
summary(percell_lmTOGET_unweightedMPD_taxalab)


plot_unweightedMPD_percell <- ggplot(filter(unweighted_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y =log10(fracprod_per_cell_noinf), color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Unweighted Mean Pairwise Dist") + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm") +
  theme(legend.position = "none", legend.title = element_blank()) +
  annotate("text", x = -3, y=-6, 
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(percell_lmPA_unweightedMPD_taxalab)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(percell_lmPA_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 3))) +
  annotate("text", x = -3, y=-7, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(percell_lmFL_unweightedMPD_taxalab)$adj.r.squared, digits = 2), "\n", 
                       "p =", round(unname(summary(percell_lmFL_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 3))) +
    annotate("text", x = -2.8, y=-8, 
         color = "black", fontface = "bold",
         label = paste("Combined: R2 =", round(summary(percell_lmTOGET_unweightedMPD_taxalab)$adj.r.squared, digits = 2), "\n", 
                       "Combined: p =", round(unname(summary(percell_lmTOGET_unweightedMPD_taxalab)$coefficients[,4][2]), digits = 4)))


############ WEIGHTED MPD
##########  MPD ANALYSIS 
# Free-Living Model
percell_lmFL_weightedMPD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction == "WholeFree"))
summary(percell_lmFL_weightedMPD_taxalab)

# Particle-associated model
percell_lmPA_weightedMPD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction == "WholePart"))
summary(percell_lmPA_weightedMPD_taxalab)


###  COMBINE BOTH PARTICLE AND FREE INTO ONE MODEL
percell_lmTOGET_weightedMPD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
                                      data = filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(percell_lmTOGET_weightedMPD_taxalab)

# Residual analysis of the UNWEIGHTED MPD Models: PARTICLE-ASSOCIATED
ALL_weight_sesMPD_df <- dplyr::filter(WEIGHTED_sesMPD_taxalab, 
                                      fraction %in% c("WholePart", "WholeFree") & !is.na(fracprod_per_cell_noinf))

plot_residuals(lm_model = percell_lmTOGET_weightedMPD_taxalab, 
               lm_observed_y = log10(ALL_weight_sesMPD_df$fracprod_per_cell_noinf),
               main_title = "Both Fractions Weighted MPD")



plot_weightedMPD_percell_TOGETHER <- ggplot(filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y =log10(fracprod_per_cell_noinf))) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3,  aes(color = fraction)) +
  xlab("Weighted Mean Pairwise Dist") + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm", color = "black") +
  theme(legend.position = "none") +
  annotate("text", x = -3, y=-6, 
           color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(percell_lmTOGET_weightedMPD_taxalab)$adj.r.squared, digits = 3), "\n", 
                         "p =", round(unname(summary(percell_lmTOGET_weightedMPD_taxalab)$coefficients[,4][2]), digits = 3)))+
  annotate("text", x = -3, y=-6.5, color = "#CCBB51", fontface = "bold", label = paste("Separate = NS"))



plot_weightedMPD_percell <- ggplot(filter(WEIGHTED_sesMPD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mpd.obs.z, y =log10(fracprod_per_cell_noinf), color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Weighted Mean Pairwise Dist") + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  #geom_smooth(method = "lm") +
  theme(legend.position = "none", legend.title = element_blank()) #+
  #annotate("text", x = -2, y=-6.5, 
  #         color = "firebrick3", fontface = "bold",
  #         label = paste("R2 =", round(summary(percell_lmPA_weightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                       "p =", round(unname(summary(percell_lmPA_weightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) +
  #annotate("text", x = 0, y=-8.2, 
  #       color = "cornflowerblue", fontface = "bold",
  #       label = paste("R2 =", round(summary(percell_lmFL_weightedMPD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                     "p =", round(unname(summary(percell_lmFL_weightedMPD_taxalab)$coefficients[,4][2]), digits = 4))) 





####### UNWEIGHTED MNTD ANALYSIS
# Free-living
percell_lmFL_unweightedMNTD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction == "WholeFree"))
summary(percell_lmFL_unweightedMNTD_taxalab)

# Particle-associated
percell_lmPA_unweightedMNTD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction == "WholePart"))
summary(percell_lmPA_unweightedMNTD_taxalab)

# BOTH FRACTIONS TOGETHER
percell_lmTOGET_unweightedMNTD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mntd.obs.z, 
                                      data = filter(unweighted_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(percell_lmTOGET_unweightedMNTD_taxalab)


plot_unweightedMNTD_percell <- ggplot(filter(unweighted_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mntd.obs.z, y = log10(fracprod_per_cell_noinf), color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Unweighted Mean Nearest Taxon Dist") + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  scale_color_manual(values = fraction_colors,
                 breaks=c("WholeFree", "WholePart"), 
                 labels=c("Free", "Particle")) + 
  geom_smooth(method = "lm", data = filter(unweighted_sesMNTD_taxalab, fraction == "WholePart")) +
  theme(legend.position = c(0.87, 0.87), 
        legend.title = element_blank()) +
  annotate("text", x = 1.5, y=-7.2,
           color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(percell_lmPA_unweightedMNTD_taxalab)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(percell_lmPA_unweightedMNTD_taxalab)$coefficients[,4][2]), digits = 3))) #+
  #annotate("text", x = -2.4, y=-7.8, 
  #       color = "cornflowerblue", fontface = "bold",
  #       label = paste("R2 =", round(summary(percell_lmFL_unweightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                     "p =", round(unname(summary(percell_lmFL_unweightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) 


####### WEIGHTED MPD ANALYSIS
# Free-living
percell_lmFL_weightedMNTD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholeFree"))
summary(percell_lmFL_weightedMNTD_taxalab)
# Particle-associated
percell_lmPA_weightedMNTD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholePart"))
summary(percell_lmPA_weightedMNTD_taxalab)

###  COMBINE BOTH PARTICLE AND FREE INTO ONE MODEL
percell_lmTOGET_weightedMNTD_taxalab <- lm(log10(fracprod_per_cell_noinf) ~ mntd.obs.z, 
                                      data = filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")))
summary(percell_lmTOGET_weightedMNTD_taxalab)  # NOT SIGNIFICANT 

# Residual analysis of the UNWEIGHTED MPD Models: PARTICLE-ASSOCIATED
ALL_weight_sesMNTD_df <- dplyr::filter(WEIGHTED_sesMNTD_taxalab, 
                                      fraction %in% c("WholePart", "WholeFree") & !is.na(fracprod_per_cell_noinf))


plot_weightedMNTD_percell <- ggplot(filter(WEIGHTED_sesMNTD_taxalab, fraction %in% c("WholePart", "WholeFree")), 
       aes(x = mntd.obs.z, y = log10(fracprod_per_cell_noinf), color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) +
  xlab("Weighted Mean Nearest Taxon Dist") + 
  ylab("log10(Production/Cell)\n (μgC/cell/hr)") + 
  scale_color_manual(values = c("firebrick3", "cornflowerblue")) +
  geom_smooth(method = "lm", data = filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholeFree")) +
  theme(legend.position = "none", legend.title = element_blank()) +
  #annotate("text", x = -2.5, y=-5.9,
  #         color = "firebrick3", fontface = "bold",
  #         label = paste("R2 =", round(summary(percell_lmPA_weightedMNTD_taxalab)$adj.r.squared, digits = 4), "\n", 
  #                       "p =", round(unname(summary(percell_lmPA_weightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) +
  annotate("text", x = 1.5, y=-8, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(percell_lmFL_weightedMNTD_taxalab)$adj.r.squared, digits = 2), "\n", 
                       "p =", round(unname(summary(percell_lmFL_weightedMNTD_taxalab)$coefficients[,4][2]), digits = 4))) 






plot_grid(plot_unweightedMPD_percell, plot_unweightedMNTD_percell,
          plot_weightedMPD_percell, plot_weightedMNTD_percell,
  labels = c("A", "B", "C", "D"),
  nrow= 1, ncol = 4
)

```


## Differences in slopes and interaction terms?

### Unweighted MPD slopes
```{r post-hoc-unweightedMPD, dependson = c("taxalab-production-BEF", "taxa-labels"),fig.width = 12, fig.height = 4,}
# UNWEIGHTED MPD 
# Are the slopes of the fraction linear models different from each other in predicting fraction production?
unweightMPD_fraction_dat <- filter(unweighted_sesMPD_taxalab, 
                       fraction %in% c("WholePart", "WholeFree")) %>%
  dplyr::select(norep_filter_name, mpd.obs.z, fraction, frac_bacprod, fracprod_per_cell_noinf) 

# Help from:  http://r-eco-evo.blogspot.com/2011/08/comparing-two-regression-slopes-by.html
# WITH an interaction term 
lm2_bulk_interaction <- lm(frac_bacprod ~ mpd.obs.z * fraction, data = unweightMPD_fraction_dat)
summary(lm2_bulk_interaction)

# Anova was run from this tutorial: http://rcompanion.org/rcompanion/e_04.html
Anova(lm2_bulk_interaction, type="II") # Since the interaction is significant, the slops across the groups *ARE* different 

# Check the residuals
plot_residuals(lm_model = lm2_bulk_interaction, 
               lm_observed_y = unweightMPD_fraction_dat$frac_bacprod,
               main_title = "Unweighted MPD with Interaction Term")

# WITHOUT an interaction term 
lm2_bulk <- lm(frac_bacprod ~ mpd.obs.z + fraction, data = unweightMPD_fraction_dat)
summary(lm2_bulk)

# Anova was run from this tutorial: http://rcompanion.org/rcompanion/e_04.html
Anova(lm2_bulk, type="II") # Since the interaction is significant, the slope across the groups *ARE* different 

# Check the residuals
plot_residuals(lm_model = lm2_bulk, 
               lm_observed_y = unweightMPD_fraction_dat$frac_bacprod,
               main_title = "Unweighted MPD with Interaction Term")

# Does the interaction term include important information not accounted for already?
Anova(lm2_bulk_interaction, lm2_bulk)

# Separate Linear Models with subsetted data
summary(lm(frac_bacprod ~ mpd.obs.z, data = dplyr::filter(unweightMPD_fraction_dat,fraction == "WholePart")))
summary(lm(frac_bacprod ~ mpd.obs.z, data = dplyr::filter(unweightMPD_fraction_dat,fraction == "WholeFree")))


## PER CELL
lm_unweightMPD_by_fraction_percell <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z/fraction, data = unweightMPD_fraction_dat)
summary(lm_unweightMPD_by_fraction_percell)

# Help from:  http://r-eco-evo.blogspot.com/2011/08/comparing-two-regression-slopes-by.html
# WITH an interaction term 
lm2_percell_interaction <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z * fraction, data = unweightMPD_fraction_dat)
summary(lm2_percell_interaction)

# Anova was run from this tutorial: http://rcompanion.org/rcompanion/e_04.html
Anova(lm2_percell_interaction, type="II") # Since the interaction is NOT significant, the slopes across the groups *ARE NOT* different!

# Check the residuals
plot_residuals(lm_model = lm2_percell_interaction, 
               lm_observed_y = unweightMPD_fraction_dat$log10(fracprod_per_cell_noinf) ,
               main_title = "Unweighted MPD with Interaction Term")

# WITHOUT an interaction term 
lm2_bulk <- lm(frac_bacprod ~ mpd.obs.z + fraction, data = unweightMPD_fraction_dat)
summary(lm2_bulk)
# Check the residuals
plot_residuals(lm_model = lm2_bulk, 
               lm_observed_y = unweightMPD_fraction_dat$frac_bacprod,
               main_title = "Unweighted MPD with Interaction Term")





summary(lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
           data = dplyr::filter(unweightMPD_fraction_dat,fraction == "WholePart")))
summary(lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
           data = dplyr::filter(unweightMPD_fraction_dat,fraction == "WholeFree")))

mod1_percell <- aov(fracprod_per_cell_noinf ~ mpd.obs.z*fraction, data = unweightMPD_fraction_dat)
summary(mod1_percell) # With interaction term 
mod2_percell <- aov(fracprod_per_cell_noinf ~ mpd.obs.z+fraction, data = unweightMPD_fraction_dat)
summary(mod2_percell) # Without interaction term 

anova(mod1_percell, mod2_percell) # Does the interaction term have a significant effect?
```


### Weighted MPD slopes
```{r post-hoc-WEIGHTEDMPD-bulk, dependson = c("taxalab-production-BEF", "taxa-labels")}
####  WEIGHTED MPD
weightMPD_fraction_dat <- filter(WEIGHTED_sesMPD_taxalab, 
                       fraction %in% c("WholePart", "WholeFree")) %>%
  dplyr::select(norep_filter_name, mpd.obs.z, fraction, frac_bacprod, fracprod_per_cell_noinf) 


# Double check values from above models
lm_WEIGHTED_MPD_by_fraction <- lm(frac_bacprod ~ mpd.obs.z/fraction, data = weightMPD_fraction_dat)
summary(lm_WEIGHTED_MPD_by_fraction)

summary(lm(frac_bacprod ~ mpd.obs.z, data = dplyr::filter(weightMPD_fraction_dat,fraction == "WholePart")))
summary(lm(frac_bacprod ~ mpd.obs.z, data = dplyr::filter(weightMPD_fraction_dat,fraction == "WholeFree")))

# Help from:  http://r-eco-evo.blogspot.com/2011/08/comparing-two-regression-slopes-by.html
mod1_bulk <- lm(frac_bacprod ~ mpd.obs.z*fraction, data = weightMPD_fraction_dat)
summary(mod1_bulk) # With interaction term 



mod2_bulk <- lm(frac_bacprod ~ mpd.obs.z+fraction, data = weightMPD_fraction_dat)
summary(mod2_bulk) # Without interaction term 

anova(mod1_bulk, mod2_bulk) # Does the interaction term have a significant effect?


## PER CELL
lm_WEIGHTED_MPD_by_fraction_percell <- lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z/fraction, data = weightMPD_fraction_dat)
summary(lm_WEIGHTED_MPD_by_fraction_percell)

summary(lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
           data = dplyr::filter(weightMPD_fraction_dat,fraction == "WholePart")))
summary(lm(log10(fracprod_per_cell_noinf) ~ mpd.obs.z, 
           data = dplyr::filter(weightMPD_fraction_dat,fraction == "WholeFree")))

mod1_percell <- lm(fracprod_per_cell_noinf ~ mpd.obs.z*fraction, data = weightMPD_fraction_dat)
summary(mod1_percell) # With interaction term 
mod2_percell <- lm(fracprod_per_cell_noinf ~ mpd.obs.z+fraction, data = weightMPD_fraction_dat)
summary(mod2_percell) # Without interaction term 

anova(mod1_percell, mod2_percell) # Does the interaction term have a significant effect?
```


# Figure 3
```{r fig-3, fig.width= 16, fig.height = 8, dependson=c("taxalab-comparison", "taxalab-percell")}
plot_grid(taxlab_unweight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums1$b)-0.5), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS")) +
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_unweight_mntd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mntd_nums1$b)-0.5), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS")) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_weight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums2$b)-1), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MPD_wilcox$p.value, digits = 6))) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_weight_mntd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mntd_nums2$b)+1.2), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MNTD_wilcox$p.value, digits = 3))) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          plot_unweightedMPD_percell + xlim(-4.5, 3.5), 
          plot_unweightedMNTD_percell + xlim(-4.5, 3.5), 
          plot_weightedMPD_percell + xlim(-4.5, 3.5), 
          plot_weightedMNTD_percell + xlim(-4.5, 3.5), 
          labels = c("A", "B", "C", "D", "E", "F", "G", "H"),
          ncol = 4, nrow = 2)



###  COMBINE TOGETHER FRACTIONS FOR FIGURE G
plot_grid(taxlab_unweight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums1$b)-0.5), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS")) +
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_unweight_mntd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mntd_nums1$b)-0.5), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS")) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_weight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums2$b)-1), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MPD_wilcox$p.value, digits = 6))) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_weight_mntd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mntd_nums2$b)+1.2), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MNTD_wilcox$p.value, digits = 3))) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          plot_unweightedMPD_percell + xlim(-4.5, 3.5), 
          plot_unweightedMNTD_percell + xlim(-4.5, 3.5), 
          plot_weightedMPD_percell_TOGETHER + xlim(-4.5, 3.5), 
          plot_weightedMNTD_percell + xlim(-4.5, 3.5), 
          labels = c("A", "B", "C", "D", "E", "F", "G", "H"),
          ncol = 4, nrow = 3)




```


```{r super-fig3, fig.width= 16, fig.height = 12, dependson=c("taxalab-comparison", "taxalab-percell", "taxalab-production-BEF")}
plot_grid(taxlab_unweight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums1$b)-0.5), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS")) +
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_unweight_mntd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mntd_nums1$b)-0.5), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS")) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_weight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums2$b)-1), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MPD_wilcox$p.value, digits = 6))) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_weight_mntd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mntd_nums2$b)+1.2), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MNTD_wilcox$p.value, digits = 3))) + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          plot_unweightedMPD_prod, 
          plot_unweightedMNTD_prod,
          plot_weightedMPD_prod, 
          plot_weightedMNTD_prod,
          plot_unweightedMPD_percell + xlim(-4.5, 3.5), 
          plot_unweightedMNTD_percell + xlim(-4.5, 3.5), 
          plot_weightedMPD_percell_TOGETHER + xlim(-4.5, 3.5), 
          plot_weightedMNTD_percell + xlim(-4.5, 3.5), 
          labels = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L"),
          ncol = 4, nrow = 3)


```


```{r fig3-subsetted, fig.width= 8.5, fig.height = 10, dependson=c("taxalab-comparison", "taxalab-percell", "taxalab-production-BEF")}
plot_grid(taxlab_unweight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums1$b)-0.5), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("NS")) +
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          taxlab_weight_mpd + ylim(-4.5, 3.5) + xlab("") + coord_flip() + 
            annotate("text", x=1.55, y=(max(mpd_nums2$b)-1), fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MPD_wilcox$p.value, digits = 6)))  + 
            theme(axis.text.y = element_text(angle=90, hjust=0.5)), 
          plot_unweightedMPD_prod, 
          plot_weightedMPD_prod_TOGETHER, 
          plot_unweightedMPD_percell + xlim(-4.5, 3.5),  
          plot_weightedMPD_percell_TOGETHER + xlim(-4.5, 3.5),
          labels = c("A", "B", "C", "D", "E", "F"),
          ncol = 2, nrow = 3)


```



# Phylogenetic Diversity vs Richness/Evenness

```{r diversity-vs-diversity, fig.width = 8, fig.height = 4, dependson = c("subset-diversity","load-data","vean-otu-alphadiv-calc", "taxa-labels")}
### Richness DF
combined_rich_unweightedMPD <- ML_otu_rich_stats %>%
  dplyr::select(norep_filter_name, mean, sd, measure) %>%
  dplyr::left_join(unweighted_sesMPD_taxalab, by = "norep_filter_name") 

summary(lm(mean ~ mpd.obs.z, data = dplyr::filter(combined_rich_unweightedMPD, fraction == "WholePart")))
summary(lm(mean ~ mpd.obs.z, data = dplyr::filter(combined_rich_unweightedMPD, fraction == "WholeFree")))
summary(lm(mean ~ mpd.obs.z, data = combined_rich_unweightedMPD))

divs_p1 <- ggplot(combined_rich_unweightedMPD, aes(y = mean, x = mpd.obs.z)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3, aes(color = fraction)) + 
  scale_color_manual(values = fraction_colors) + 
  geom_smooth(method = "lm", color = "black") +
  xlab("Unweighted MPD") + ylab("Observed Richness") +
  theme(legend.position = c(0.85, 0.9),
        legend.title = element_blank()) +
  annotate("text", x = -3, y=200, 
         color = "black", fontface = "bold",
         label = paste("R2 =", round(summary(lm(mean ~ mpd.obs.z, data = combined_rich_unweightedMPD))$adj.r.squared, 
                                     digits = 2), "\n", 
                       "p =", round(unname(summary(lm(mean ~ mpd.obs.z, data = combined_rich_unweightedMPD))$coefficients[,4][2]), 
                                    digits = 4))) 

# Plot both of the models 
divs_p2 <- ggplot(combined_rich_unweightedMPD, aes(y = mean, x = mpd.obs.z, color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) + 
  scale_color_manual(values = fraction_colors) + 
  geom_smooth(method = "lm", data = dplyr::filter(combined_rich_unweightedMPD, fraction == "WholePart")) +
  xlab("Unweighted MPD") + ylab("Observed Richness") +
  theme(legend.position = c(0.85, 0.9),
        legend.title = element_blank()) +
  annotate("text", x = -3, y=400, 
         color = "firebrick3", fontface = "bold",
         label = paste("R2 =", round(summary(lm(mean ~ mpd.obs.z, 
                                                data = dplyr::filter(combined_rich_unweightedMPD, fraction == "WholePart")))$adj.r.squared, 
                                     digits = 2), "\n", 
                       "p =", round(unname(summary(lm(mean ~ mpd.obs.z, 
                                                      data = dplyr::filter(combined_rich_unweightedMPD, fraction == "WholePart")))$coefficients[,4][2]), 
                                    digits = 3))) +
  annotate("text", x = -3, y=200, 
         color = "cornflowerblue", fontface = "bold",
         label = paste("R2 =", round(summary(lm(mean ~ mpd.obs.z, 
                                                data = dplyr::filter(combined_rich_unweightedMPD, fraction == "WholeFree")))$adj.r.squared, 
                                     digits = 2), "\n", 
                       "p =", round(unname(summary(lm(mean ~ mpd.obs.z, 
                                                      data = dplyr::filter(combined_rich_unweightedMPD, fraction == "WholeFree")))$coefficients[,4][2]), 
                                    digits = 3))) 




# Try it with SES MNTD
combined_rich_unweightedMNTD <- ML_otu_rich_stats %>%
  dplyr::select(norep_filter_name, mean, sd, measure) %>%
  dplyr::left_join(unweighted_sesMNTD_taxalab, by = "norep_filter_name") 

summary(lm(mean ~ mntd.obs.z, data = dplyr::filter(combined_rich_unweightedMNTD, fraction == "WholePart")))  # NS
summary(lm(mean ~ mntd.obs.z, data = dplyr::filter(combined_rich_unweightedMNTD, fraction == "WholeFree"))) # 
summary(lm(mean ~ mntd.obs.z, data = combined_rich_unweightedMNTD))

ggplot(combined_rich_unweightedMNTD, aes(y = mean, x = mntd.obs.z)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3, aes(color = fraction)) + 
  scale_color_manual(values = fraction_colors) + 
  geom_smooth(method = "lm", data = dplyr::filter(combined_rich_unweightedMNTD, fraction == "WholeFree"), color = "cornflowerblue") +
  xlab("Unweighted MNTD") + ylab("Observed Richness") +
  theme(legend.position = c(0.85, 0.85),
        legend.title = element_blank()) +
  annotate("text", x = 2, y=200, 
       color = "cornflowerblue", fontface = "bold",
       label = paste("R2 =", round(summary(lm(mean ~ mntd.obs.z, 
                                              data = dplyr::filter(combined_rich_unweightedMNTD, fraction == "WholeFree")))$adj.r.squared, 
                                   digits = 2), "\n", 
                     "p =", round(unname(summary(lm(mean ~ mntd.obs.z, 
                                                    data = dplyr::filter(combined_rich_unweightedMNTD, fraction == "WholeFree")))$coefficients[,4][2]), 
                                  digits = 3))) 





### INVERSE SIMPS
combined_invsimps_weightedMPD <- ML_otu_invsimps_stats %>%
  dplyr::select(norep_filter_name, mean, sd, measure) %>%
  dplyr::left_join(WEIGHTED_sesMPD_taxalab, by = "norep_filter_name") 

summary(lm(mean ~ mpd.obs.z, data = dplyr::filter(combined_invsimps_weightedMPD, fraction == "WholePart")))
summary(lm(mean ~ mpd.obs.z, data = dplyr::filter(combined_invsimps_weightedMPD, fraction == "WholeFree")))
summary(lm(mean ~ mpd.obs.z, data = combined_invsimps_weightedMPD))


# All points together
evendivs_p1 <- ggplot(combined_invsimps_weightedMPD, aes(y = mean, x = mpd.obs.z)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3, aes( color = fraction)) + 
  scale_color_manual(values = fraction_colors) + 
  geom_smooth(color = "black") +
  xlab("Weighted MPD") + ylab("Inverse Simpson") +
  theme(legend.position = c(0.85, 0.85),
        legend.title = element_blank())

# Models separately
evendivs_p2 <- ggplot(combined_invsimps_weightedMPD, aes(y = mean, x = mpd.obs.z, color = fraction)) + 
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5) + xlim(-4.5, 3.5) +
  geom_point(size = 3) + 
  scale_color_manual(values = fraction_colors) + 
  geom_smooth(method = "lm") +
  xlab("Weighted MPD") + ylab("Inverse Simpson") +
  theme(legend.position = c(0.85, 0.85),
        legend.title = element_blank()) +
  annotate("text", x = -3, y=70, 
     color = "cornflowerblue", fontface = "bold",
     label = paste("R2 =", round(summary(lm(mean ~ mpd.obs.z, 
                                            data = dplyr::filter(combined_invsimps_weightedMPD, fraction == "WholeFree")))$adj.r.squared, 
                                 digits = 2), "\n", 
                   "p =", round(unname(summary(lm(mean ~ mpd.obs.z, 
                                                  data = dplyr::filter(combined_invsimps_weightedMPD, fraction == "WholeFree")))$coefficients[,4][2]), 
                                digits = 4))) +
    annotate("text", x = 2.6, y=48, 
     color = "firebrick3", fontface = "bold",
     label = paste("R2 =", round(summary(lm(mean ~ mpd.obs.z, 
                                            data = dplyr::filter(combined_invsimps_weightedMPD, fraction == "WholePart")))$adj.r.squared, 
                                 digits = 2), "\n", 
                   "p =", round(unname(summary(lm(mean ~ mpd.obs.z, 
                                                  data = dplyr::filter(combined_invsimps_weightedMPD, fraction == "WholePart")))$coefficients[,4][2]), 
                                digits = 3))) 

####  WEIGHTED MNTD: NEAREST TAXON
combined_invsimps_weightedMNTD <- ML_otu_invsimps_stats %>%
  dplyr::select(norep_filter_name, mean, sd, measure) %>%
  dplyr::left_join(WEIGHTED_sesMNTD_taxalab, by = "norep_filter_name") 

summary(lm(mean ~ mntd.obs.z, data = dplyr::filter(combined_invsimps_weightedMNTD, fraction == "WholePart"))) # NS
summary(lm(mean ~ mntd.obs.z, data = dplyr::filter(combined_invsimps_weightedMNTD, fraction == "WholeFree"))) # NS
summary(lm(mean ~ mntd.obs.z, data = combined_invsimps_weightedMNTD)) # NS




### SIMPS EVENESS
combined_simpseven_weightedMPD <- ML_otu_simpseven_stats %>%
  dplyr::select(norep_filter_name, mean, sd, measure) %>%
  dplyr::left_join(WEIGHTED_sesMPD_taxalab, by = "norep_filter_name") 
# Very similar to inverse simpson and MPD 
summary(lm(mean ~ mpd.obs.z, data = dplyr::filter(combined_simpseven_weightedMPD, fraction == "WholePart"))) # R2 = 0.25, pval = 0.058
summary(lm(mean ~ mpd.obs.z, data = dplyr::filter(combined_simpseven_weightedMPD, fraction == "WholeFree"))) # R2 = 0.64, pval = 0.001
summary(lm(mean ~ mpd.obs.z, data = combined_simpseven_weightedMPD)) # NS

```

```{r figure-4, fig.width = 7, fig.height = 6, dependson = c("subset-diversity","load-data","vean-otu-alphadiv-calc", "taxa-labels", "diversity-vs-diversity")}
# PLOT
plot_grid(divs_p1, divs_p2, # Richness vs UNweighted MPD
          evendivs_p1, evendivs_p2, # Inverse Simpson vs Weighted MPD 
          nrow = 2, ncol = 2,
          labels = c("A", "B", "C", "D"))
```

## Residual Analysis 
```{r check-lm-phylogenetic-residuals, fig.width = 12, fig.height = 4, dependson = "plot-vegan-fracprod"}
##########################################################################
#############################   RESIDUALS   ##############################
##########################################################################

######################################################### UNWEIGHTED MPD Models
# Residual analysis of the UNWEIGHTED MPD Models: FREE-LIVING
FL_unweight_sesMPD_df <- dplyr::filter(unweighted_sesMPD_taxalab, fraction == "WholeFree")

plot_residuals(lm_model = percell_lmFL_unweightedMPD_taxalab, 
               lm_observed_y = log10(FL_unweight_sesMPD_df$fracprod_per_cell_noinf),
               main_title = "Free-Living Unweighted MPD")

# Residual analysis of the UNWEIGHTED MPD Models: PARTICLE-ASSOCIATED
PA_unweight_sesMPD_df <- dplyr::filter(unweighted_sesMPD_taxalab, fraction == "WholePart" & !is.na(fracprod_per_cell_noinf))

plot_residuals(lm_model = percell_lmPA_unweightedMPD_taxalab, 
               lm_observed_y = log10(PA_unweight_sesMPD_df$fracprod_per_cell_noinf),
               main_title = "Particle-Associated Unweighted MPD")



######################################################### UNWEIGHTED MNTD
# Residual analysis of the UNWEIGHTED MNTD Models: PARTICLE-ASSOCIATED
PA_unweight_sesMNTD_df <- filter(unweighted_sesMNTD_taxalab, fraction == "WholePart" & !is.na(fracprod_per_cell_noinf))

plot_residuals(lm_model = percell_lmPA_unweightedMNTD_taxalab, 
               lm_observed_y = log10(PA_unweight_sesMNTD_df$fracprod_per_cell_noinf),
               main_title = "Particle-Associated Unweighted MNTD")


######################################################### WEIGHTED MNTD 
# Residual analysis of the WEIGHTED MNTD Models: FREE-LIVING
FL_unweight_sesMNTD_df <- filter(WEIGHTED_sesMNTD_taxalab, fraction == "WholeFree" & !is.na(fracprod_per_cell_noinf))

plot_residuals(lm_model = percell_lmFL_weightedMNTD_taxalab, 
               lm_observed_y = log10(FL_unweight_sesMNTD_df$fracprod_per_cell_noinf),
               main_title = "Free-Living Weighted MNTD")
```


```{r old-fig3, fig.width = 8.5, fig.height = 7, dependson = c("taxa-labels","taxalab-vs-fracprod", "taxalab-comparison")}
# Plot it altogether 
plot_grid(unweightedMPD_vs_fracprod_taxlab + ggtitle("") +
            xlab("Unweighted Mean Pairwise Distance") +
            theme(legend.position = "none"), 
          unweightedMNTD_vs_fracprod_taxlab + ggtitle("") +
            xlab("Unweighted Mean Nearest Taxon Distance") +
            theme(legend.position = "none"), 
          taxlab_weight_mpd +  xlab("\n Fraction \n") +
            theme(axis.text.y = element_blank()) + 
            ylab("Weighted Mean Pairwise Distance") +
            annotate("text", x=1.55, y=1.6, fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MPD_wilcox$p.value, digits = 6))) + 
            coord_flip(),
          taxlab_weight_mntd + xlab("\n Fraction \n") +
            theme(axis.text.y = element_blank()) + 
            ylab("Weighted Mean Nearest Taxon Distance") +
            annotate("text", x=1.5, y=0.5, fontface = "bold",  size = 3.5, color = "gray40",
                       label= paste("***\np =", round(weight_MNTD_wilcox$p.value, digits = 3))) + 
            coord_flip(), 
          labels = c("A", "B", "C", "D"), 
          ncol = 2, nrow = 2)
```

## Congruency between fractions 
```{r frac-congruency, fig.width=4, fig.height=4, dependson="subset-diversity"}
rich_df <- ML_otu_rich_stats %>%
  dplyr::select(norep_filter_name, mean, sd, measure, fraction) %>%
  mutate(norep_water_name = paste(substr(norep_filter_name, 1,4), substr(norep_filter_name, 6,9), sep = "")) %>%
  spread(fraction, mean) 

rich_part_df <- dplyr::select(rich_df, norep_water_name, WholePart) %>%
  distinct() %>%  filter(!is.na(WholePart))

rich_free_df <- dplyr::select(rich_df, norep_water_name, WholeFree) %>%
  distinct() %>%  filter(!is.na(WholeFree))

combined_rich_df <- left_join(rich_free_df, rich_part_df, by = "norep_water_name") %>%
  rename(names = norep_water_name) %>%
  make_metadata_norep() %>%
  dplyr::select(-c(year, fraction, month, season, nuc_acid_type, project))


summary(lm(WholePart ~ WholeFree, data = combined_rich_df))


ggplot(combined_rich_df, aes(x=WholePart, y=WholeFree)) + 
  geom_point(size = 3.5) +
  geom_abline(intercept = 0, slope = 1)
```




# Beta Diversity Analysis

```{r scale-reads, dependson = "load-data"}
scale_otu_merged_musk_pruned <- scale_reads(otu_merged_musk_pruned_noMOTHJ715_MBRHP715, round = "matround")
scale_otu_merged_musk_pruned # ALL Samples phyloseq object


# Check the sequencing depth of each sample 
scaled_sums_otu <- data.frame(rowSums(otu_table(scale_otu_merged_musk_pruned)))
colnames(scaled_sums_otu) <- "Sample_TotalSeqs"
scaled_sums_otu$names <- row.names(scaled_sums_otu)
scaled_sums_otu <- arrange(scaled_sums_otu, Sample_TotalSeqs) %>%
  make_metadata_norep()

##  Plot based on all samples 
plot_sample_sums(dataframe = scaled_sums_otu, x_total_seqs = "Sample_TotalSeqs", fill_variable = "project")

##  Plot based on depth of samples 
plot_sample_sums(dataframe = scaled_sums_otu, x_total_seqs = "Sample_TotalSeqs", fill_variable = "limnion")

##  Plot based on depth of samples WITHOUT SEDIMENT!
plot_sample_sums(dataframe = filter(scaled_sums_otu, limnion != "Benthic"), 
                 x_total_seqs = "Sample_TotalSeqs", fill_variable = "limnion")


### Subset out samples that we only have productivity data for
productivity_scale_int <- subset_samples(scale_otu_merged_musk_pruned, limnion == "Top" & year == "2015" & 
                                           fraction %in% c("WholePart", "WholeFree"))

productivity_scale <- prune_taxa(taxa_sums(productivity_scale_int) > 0, productivity_scale_int) 
```


```{r ordinate}
productivity_bray <- phyloseq::distance(productivity_scale, method = "bray")

# make a data frame from the sample_data
sampledf <- data.frame(sample_data(productivity_scale))

# Adonis test
adonis(productivity_bray ~ frac_bacprod, data = sampledf)
```


```{r calculate-bray-curtis, fig.width=10, fig.height=4, dependson = "scale-reads"}
scaled_productivity_otu_df <- as.matrix(otu_table(productivity_scale))

bray_productivity_otu <- vegdist(scaled_productivity_otu_df, method = "bray", binary = FALSE, upper = TRUE)

# Melt the bray curtis distance to a dataframe
bray <- reshape2::melt(as.matrix(bray_productivity_otu), varnames = c("samp1", "samp2"))
bray <- subset(bray, value > 0) # Remove the samples compared to themselves


bray$lakesite1 <- substr(bray$samp1,2,3) # Create a new column called lakenames1 with first 3 letters of string
bray$lakesite2 <- substr(bray$samp2,2,3) # Create a new column called lakenames2 with first 3 letters of string
bray$limnion1 <- substr(bray$samp1, 4, 4) # Create a column called limnon1 with hypo or epi
bray$limnion2 <- substr(bray$samp2, 4, 4) # Create a column called limnon2 with hypo or epi
bray$filter1 <- substr(bray$samp1, 5, 5)  # Create a column called filter1 with PA or FL
bray$filter2 <- substr(bray$samp2, 5, 5) # Create a column called filter2 with PA or FL
bray$month1 <- substr(bray$samp1, 6, 6)  # Create a column called month1 with may, july, or september
bray$month2 <- substr(bray$samp2, 6, 6) # Create a column called month2 with may, july, or september
bray$year1 <- substr(bray$samp1, 7, 9) # Create a column called year1 with 2014 or 2015
bray$year2 <- substr(bray$samp2, 7, 9) # Create a column called year2 with 2014 or 2015


  # Depth in water column
bray$limnion1 <- ifelse(bray$limnion1 == "E", "Top", NA)
bray$limnion2 <- ifelse(bray$limnion2 == "E", "Top", NA)
  
# fraction Fraction
bray$filter1 <- ifelse(bray$filter1 == "F", "Free", 
                             ifelse(bray$filter1 == "P", "Particle",
                                    ifelse(bray$filter1 == "J","WholePart",
                                           ifelse(bray$filter1 == "K","WholeFree", NA))))
bray$filter2 <- ifelse(bray$filter2 == "F", "Free", 
                             ifelse(bray$filter2 == "P", "Particle",
                                    ifelse(bray$filter2 == "J","WholePart",
                                           ifelse(bray$filter2 == "K","WholeFree", NA))))


# Stop calculations if sample1 is equal to sample2
stopifnot(nrow(dplyr::filter(bray, samp1 == samp2)) == 0)


##  Add productivity data to bray dataframe
prod_data <- sample_data(productivity_scale) %>%
  dplyr::select(norep_filter_name, frac_bacprod) %>%
  mutate(frac_bacprod = round(frac_bacprod, digits = 1),
         samp1 = norep_filter_name) %>%
  dplyr::select(-norep_filter_name)

# Add fraction production for sample1
bray_frac1 <- right_join(bray, prod_data, by = "samp1") %>%
  rename(frac_prod1 = frac_bacprod)

# Add fraction production for sample2
bray_final <- right_join(bray_frac1, rename(prod_data, samp2 = samp1), by = "samp2") %>%
  rename(frac_prod2 = frac_bacprod) %>%
  mutate(delta_frac_bacprod = abs(frac_prod1 - frac_prod2),
         filter_match = ifelse(filter1 == filter2, filter1, "diff_filter"))


ggplot(bray_final, aes(x = value, y = delta_frac_bacprod, color = filter_match)) +
  geom_point() +
  facet_grid(.~filter_match) + 
  xlab("Bray-Curtis Dissimilarity") + ylab("Difference Fraction Productvity") +
  geom_smooth() +
  theme(legend.position = c(0.9, 0.85),legend.title = element_blank())



#### Difference between filter comparisons?
bray_kruskal <- kruskal.test(value ~ as.factor(filter_match), data = bray_final)
bray_kruskal

library(pgirmess)
library(multcompView)

bray_kruskal_MC <- kruskalmc(bray_final$value ~ bray_final$filter_match)  ## Defaults to P < 0.05
#print(bray_prod_KW_MC)
### Time to figure out letters to represent significance in a plot
bray_test <- bray_kruskal_MC$dif.com$difference # select logical vector
names(bray_test) <- row.names(bray_kruskal_MC$dif.com) # add comparison names
# create a list with "homogenous groups" coded by letter
bray_letters <- multcompLetters(bray_test, compare="<", threshold=0.05, 
                                 Letters=c(letters, LETTERS, "."), reversed = FALSE)
###  Extract the values from the multcompLetters object
bray_sigs_dataframe <-  data.frame(as.vector(names(bray_letters$Letters)), as.vector(bray_letters$Letters))
colnames(bray_sigs_dataframe) <- c("filter_match", "siglabel")
bray_try <- left_join(bray_final, bray_sigs_dataframe, by = "filter_match")

bray_sigs <- bray_try %>%
  dplyr::select(filter_match, siglabel) %>%
  distinct()

bray_box <- ggplot(bray_try, aes(y = value, x = filter_match, color = filter_match, fill = filter_match)) +
  geom_boxplot(alpha = 0.3) + geom_jitter(size = 3) + 
  scale_y_continuous(limits = c(0.1, 0.95), breaks = c(0.1, 0.3, 0.5, 0.7, 0.9)) +
  geom_text(data = bray_sigs, aes(label = siglabel, x = filter_match, y = 0.95), size =5, color = "black") +
  ylab("Bray-Curtis Dissimilarity") + xlab("Filter Comparison") +
  ggtitle("Bray-Curtis") + 
  theme(legend.position = "none")
```


```{r soren-prod, fig.width=10, fig.height=4, dependson = "scale-reads"}
soren_productivity_otu <- vegdist(data.frame(otu_table(productivity_scale)), method = "bray", binary = TRUE)

# Melt the soren curtis distance to a dataframe
soren <- reshape2::melt(as.matrix(soren_productivity_otu), varnames = c("samp1", "samp2"))
soren <- subset(soren, value > 0) # Remove the samples compared to themselves


soren$lakesite1 <- substr(soren$samp1,2,3) # Create a new column called lakenames1 with first 3 letters of string
soren$lakesite2 <- substr(soren$samp2,2,3) # Create a new column called lakenames2 with first 3 letters of string
soren$limnion1 <- substr(soren$samp1, 4, 4) # Create a column called limnon1 with hypo or epi
soren$limnion2 <- substr(soren$samp2, 4, 4) # Create a column called limnon2 with hypo or epi
soren$filter1 <- substr(soren$samp1, 5, 5)  # Create a column called filter1 with PA or FL
soren$filter2 <- substr(soren$samp2, 5, 5) # Create a column called filter2 with PA or FL
soren$month1 <- substr(soren$samp1, 6, 6)  # Create a column called month1 with may, july, or september
soren$month2 <- substr(soren$samp2, 6, 6) # Create a column called month2 with may, july, or september
soren$year1 <- substr(soren$samp1, 7, 9) # Create a column called year1 with 2014 or 2015
soren$year2 <- substr(soren$samp2, 7, 9) # Create a column called year2 with 2014 or 2015


  # Depth in water column
soren$limnion1 <- ifelse(soren$limnion1 == "E", "Top", NA)
soren$limnion2 <- ifelse(soren$limnion2 == "E", "Top", NA)
  
# fraction Fraction
soren$filter1 <- ifelse(soren$filter1 == "F", "Free", 
                             ifelse(soren$filter1 == "P", "Particle",
                                    ifelse(soren$filter1 == "J","WholePart",
                                           ifelse(soren$filter1 == "K","WholeFree", NA))))
soren$filter2 <- ifelse(soren$filter2 == "F", "Free", 
                             ifelse(soren$filter2 == "P", "Particle",
                                    ifelse(soren$filter2 == "J","WholePart",
                                           ifelse(soren$filter2 == "K","WholeFree", NA))))


# Stop calculations if sample1 is equal to sample2
stopifnot(nrow(dplyr::filter(soren, samp1 == samp2)) == 0)


##  Add productivity data to soren dataframe
prod_data <- sample_data(productivity_scale) %>%
  dplyr::select(norep_filter_name, frac_bacprod) %>%
  mutate(frac_bacprod = round(frac_bacprod, digits = 1),
         samp1 = norep_filter_name) %>%
  dplyr::select(-norep_filter_name)

# Add fraction production for sample1
soren_frac1 <- right_join(soren, prod_data, by = "samp1") %>%
  rename(frac_prod1 = frac_bacprod)

# Add fraction production for sample2
soren_final <- right_join(soren_frac1, rename(prod_data, samp2 = samp1), by = "samp2") %>%
  rename(frac_prod2 = frac_bacprod) %>%
  mutate(delta_frac_bacprod = abs(frac_prod1 - frac_prod2),
         filter_match = ifelse(filter1 == filter2, filter1, "diff_filter"))


ggplot(soren_final, aes(x = value, y = delta_frac_bacprod, color = filter_match)) +
  geom_point() +
  facet_grid(.~filter_match) + 
  xlab("Sørensen Dissimilarity") + ylab("Difference Fraction Productvity") +
  geom_smooth() +
  theme(legend.position = c(0.9, 0.85),legend.title = element_blank())



###  Is there a difference in Sorensen Values between the different filter comparisons?
soren_kruskal <- kruskal.test(value ~ as.factor(filter_match), data = soren_final)
soren_kruskal

soren_kruskal_MC <- kruskalmc(soren_final$value ~ soren_final$filter_match)  ## Defaults to P < 0.05
#print(soren_prod_KW_MC)
### Time to figure out letters to represent significance in a plot
soren_test <- soren_kruskal_MC$dif.com$difference # select logical vector
names(soren_test) <- row.names(soren_kruskal_MC$dif.com) # add comparison names
# create a list with "homogenous groups" coded by letter
soren_letters <- multcompLetters(soren_test, compare="<", threshold=0.05, 
                                 Letters=c(letters, LETTERS, "."), reversed = FALSE)
###  Extract the values from the multcompLetters object
soren_sigs_dataframe <-  data.frame(as.vector(names(soren_letters$Letters)), as.vector(soren_letters$Letters))
colnames(soren_sigs_dataframe) <- c("filter_match", "siglabel")
soren_try <- left_join(soren_final, soren_sigs_dataframe, by = "filter_match")

soren_sigs <- soren_try %>%
  dplyr::select(filter_match, siglabel) %>%
  distinct()

soren_box <- ggplot(soren_try, aes(y = value, x = filter_match, color = filter_match, fill = filter_match)) +
  geom_boxplot(alpha = 0.3) + geom_jitter(size = 3) + 
  scale_y_continuous(limits = c(0.1, 0.95), breaks = c(0.1, 0.3, 0.5, 0.7, 0.9)) +
  geom_text(data = soren_sigs, aes(label = siglabel, x = filter_match, y = 0.85), size =5, color = "black") +
  ylab("Sørensen Dissimilarity") + xlab("Filter Comparison") +
  ggtitle("Sørensen") + 
  theme(legend.position = "none"); 



soren_box <- ggplot(soren_try, aes(y = value, x = filter_match, color = filter_match, fill = filter_match)) +
  geom_boxplot(alpha = 0.3) + geom_jitter(size = 3) + 
  scale_y_continuous(limits = c(0.1, 0.95), breaks = c(0.1, 0.3, 0.5, 0.7, 0.9)) +
  geom_text(data = soren_sigs, aes(label = siglabel, x = filter_match, y = 0.85), size =5, color = "black") +
  ylab("Sørensen Dissimilarity") + xlab("Filter Comparison") +
  ggtitle("Sørensen") + 
  theme(legend.position = "none")
```



```{r beta-box, dependson = c("calculate-bray-curtis","soren-prod"), fig.width=8, fig.height = 4}
plot_grid(soren_box, bray_box, labels = c("A", "B"), ncol = 2)
```
