---
title: "Rarefied Diversity Analysis"
author: "Marian L. Schmidt"
date: "February 14, 2017"
output:
  html_document:
    code_folding: show
    highlight: haddock
    keep_md: yes
    theme: united
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
---


```{r load-libraries, message = FALSE, warning = FALSE}
library(ggplot2)
library(devtools)
library(phyloseq)
library(tidyr)
library(dplyr)
library(cowplot)
library(picante) # Will also include ape and vegan 
source("Muskegon_functions.R")
source("set_colors.R")
```


# Rarefied Alpha Diversity Analysis
## Remove samples with too few reads
```{r load-data, eval = TRUE}
# Loads a phyloseq object named otu_merged_musk_pruned)
load("../data/otu_merged_musk_pruned.RData")
# The name of the phyloseq object is: 
otu_merged_musk_pruned 

# Remove tree because it's too computationally intensive
otu_merged_musk_pruned <- merge_phyloseq(tax_table(otu_merged_musk_pruned), sample_data(otu_merged_musk_pruned), otu_table(otu_merged_musk_pruned))
otu_merged_musk_pruned

# Remove MOTHJ715 and MBRHP715
otu_merged_musk_pruned_noMOTHJ715_MBRHP715 <- subset_samples(otu_merged_musk_pruned, 
                                                             norep_filter_name != "MOTHJ715" & 
                                                               norep_filter_name != "MBRHP715")
otu_merged_musk_pruned_noMOTHJ715_MBRHP715
```


```{r seq-read-depth, eval = TRUE}
# Check the sequencing depth of each sample 
sums_otu <- data.frame(rowSums(otu_table(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)))
colnames(sums_otu) <- "Sample_TotalSeqs"
sums_otu$names <- row.names(sums_otu)
sums_otu <- arrange(sums_otu, Sample_TotalSeqs) 
sums_otu <- make_metadata_norep(sums_otu)

## Add total sequences to metadata frame 
metdf <- sample_data(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
sums_otu$norep_filter_name <- sums_otu$names
sums_otu2 <- select(sums_otu, norep_filter_name, Sample_TotalSeqs)
metdf_num2 <- left_join(metdf, sums_otu2, by = "norep_filter_name")
row.names(metdf_num2) <- metdf_num2$norep_filter_name
# Rename the sample data 
sample_data(otu_merged_musk_pruned_noMOTHJ715_MBRHP715) <- metdf_num2

# Calculate Hill diversity Simpson's Evenness
meta_data <- data.frame(sample_data(otu_merged_musk_pruned_noMOTHJ715_MBRHP715))  %>%
  mutate(simps_even = D2/D0)

### PREPARE DATA FRAMES FOR PHENOFLOW ALPHA DIVERSITY ANALYSIS
free_meta_data <- filter(meta_data, fraction %in% c("Free", "WholeFree") & norep_filter_name != "MOTHJ715" & limnion == "Top")
part_meta_data <- filter(meta_data, fraction %in% c("Particle", "WholePart") & norep_filter_name != "MOTHJ715" & limnion == "Top")
nosed_meta_data <- filter(meta_data, fraction != "Sediment" & limnion == "Top")

# Only "True Free Living and Particle" From 20um prefiltered samples 
free_only <- filter(free_meta_data, fraction == "Free")
part_only <- filter(part_meta_data, fraction == "Particle")

## 2015 specific samples that have NOT been prefiltered (whole water)
wholefree_only <- filter(free_meta_data, fraction == "WholeFree")
wholepart_only <- filter(part_meta_data, fraction == "WholePart")
```


#### Rarefy Read Depth Analysis ran on December 16th, 2016
```{r calculate-alphadiv, eval = FALSE, echo = FALSE}
 # Initialize matrices to store richness and evenness estimates
# nsamp <- nsamples(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
# min_lib <- min(sample_sums(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)) - 1
# 
# otu_richness <- matrix(nrow = nsamp, ncol = 100)
# row.names(otu_richness) <- sample_names(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
# 
# otu_evenness <- matrix(nrow = nsamp, ncol = 100)
# row.names(otu_evenness) <- sample_names(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
# 
# otu_shannon <- matrix(nrow = nsamp, ncol = 100)
# row.names(otu_shannon) <- sample_names(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
## 
 ## It is always important to set a seed when you subsample so your result is replicable
 #set.seed(777)

# for (i in 1:100) {
#  # Subsample
#  r <- rarefy_even_depth(otu_merged_musk_pruned_noMOTHJ715_MBRHP715, sample.size = min_lib, verbose = FALSE, replace = TRUE)
# 
#  # Calculate richness
#  rich <- as.numeric(as.matrix(estimate_richness(r, measures = "Observed")))
#  otu_richness[ ,i] <- rich
# 
#  # Calculate evenness
#  even <- as.numeric(as.matrix(estimate_richness(r, measures = "InvSimpson")))
#  otu_evenness[ ,i] <- even
# 
#  # Calculate evenness
#  shannon <- as.numeric(as.matrix(estimate_richness(r, measures = "Shannon")))
#  otu_shannon[ ,i] <- shannon
# 
# }

#  write.table(otu_evenness, "../data/metadata/otu_evenness100_rarefy2895", row.names = TRUE)
#  write.table(otu_richness, "../data/metadata/otu_richness100_rarefy2895", row.names = TRUE)
#  write.table(otu_shannon, "../data/metadata/otu_shannon100_rarefy2895", row.names = TRUE)
```

# Fraction Diversity-Production Analysis 
```{r WHOLE-vegan-otu-alphadiv, eval = TRUE, fig.width = 10, fig.height = 8, warning = FALSE}
# Load values
nsamp <- nsamples(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)
min_lib <- min(sample_sums(otu_merged_musk_pruned_noMOTHJ715_MBRHP715)) - 1
min_lib

# Read in the files 
otu_richness <- read.table("../data/metadata/otu_richness100_rarefy2895",  header = TRUE)
otu_evenness <- read.table("../data/metadata/otu_evenness100_rarefy2895", header = TRUE)
otu_shannon <- read.table("../data/metadata/otu_shannon100_rarefy2895", header = TRUE)

# Create a new dataframe to hold the means and standard deviations of richness estimates
norep_filter_name <- row.names(otu_richness)
mean <- apply(otu_richness, 1, mean)
sd <- apply(otu_richness, 1, sd)
measure <- rep("Richness", nsamp)
otu_rich_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Create a new dataframe to hold the means and standard deviations of evenness estimates
norep_filter_name <- row.names(otu_evenness)
mean <- apply(otu_evenness, 1, mean)
sd <- apply(otu_evenness, 1, sd)
measure <- rep("Inverse_Simpson", nsamp)
otu_even_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Create a new dataframe to hold the means and standard deviations of shannon entropy estimates
norep_filter_name <- row.names(otu_shannon)
mean <- apply(otu_shannon, 1, mean)
sd <- apply(otu_shannon, 1, sd)
measure <- rep("Shannon_Entropy", nsamp)
otu_shan_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Create a new dataframe to hold the means and standard deviations of simpsons evenness estimates
norep_filter_name <- row.names(otu_evenness)
mean <- apply(otu_evenness, 1, mean)
sd <- apply(otu_evenness, 1, sd)
measure <- rep("Inverse_Simpson", nsamp)
otu_simpseven_stats <- data.frame(norep_filter_name, mean, sd, measure)

# Calculate Simpson's Evenness into new df called "simps_evenness"
otu_simps_evenness <- inner_join(otu_rich_stats, otu_even_stats, by = "norep_filter_name") %>%
  mutate(mean = mean.y/mean.x,
         sd = sd(mean),
         measure = "Simpsons_Evenness") %>%
  select(norep_filter_name, mean, sd, measure)

# Combine alpha diversity into one dataframe 
otu_alpha <- rbind(otu_rich_stats, otu_even_stats, otu_simps_evenness, otu_shan_stats)
s <- data.frame(sample_data(otu_merged_musk_pruned_noMOTHJ715_MBRHP715))
otu_alphadiv <- merge(otu_alpha, s, by = "norep_filter_name") %>%
  filter(project == "Muskegon_Lake" & limnion == "Top" & fraction != "Sediment") %>%
  dplyr::select(-D0, -D0_SD, -D0_chao, -D0_chao_sd, -D1, -D1_sd, -D2, -D2_sd)

# Set the factor levels to be the order we'd like
otu_alphadiv$fraction <- factor(otu_alphadiv$fraction,  levels = c("WholePart", "Particle", "WholeFree", "Free"))
otu_alphadiv$lakesite <- factor(otu_alphadiv$lakesite,  levels = c("MOT", "MDP", "MBR", "MIN"))
otu_alphadiv$measure <- factor(otu_alphadiv$measure,  
                               levels = c("Richness", "Simpsons_Evenness", "Shannon_Entropy", "Inverse_Simpson"))


###############################
ML_otu_rich_stats <- filter(otu_alphadiv, measure == "Richness" & project == "Muskegon_Lake" & fraction %in% c("WholePart", "WholeFree") & year == "2015")

# Can fractional production be predicted by richness? 
free_ML_otu_rich_stats <- filter(ML_otu_rich_stats, fraction == "WholeFree")
freeprod_ML_otu_rich <- lm(frac_bacprod ~ mean, data = free_ML_otu_rich_stats)
freeprod_ML_otu_rich
summary(freeprod_ML_otu_rich)

part_ML_abs_rich_stats <- filter(ML_otu_rich_stats, fraction == "WholePart")
partprod_MLotu_rich <- lm(frac_bacprod ~ mean, data = part_ML_abs_rich_stats)
partprod_MLotu_rich
summary(partprod_MLotu_rich)

# Plot 
otu_rich_vegan <- ggplot(ML_otu_rich_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ggtitle("OTU: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  #scale_x_continuous(expand = c(0,0), limits = c(0,900)) + 
  #scale_y_continuous(limits = c(0,70),expand = c(0,0)) + 
  ylab("Production (μgC/L/hr)") + xlab("Observed Richness") +
  #geom_smooth(data=subset(ML_otu_rich_stats, fraction == "Free"), method='lm') + 
  geom_smooth(data=subset(ML_otu_rich_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 500, y=45, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_rich)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_rich)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 650, y=5, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_rich)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_rich)$coefficients[,4][2]), digits = 4)));



######################################################### SHANNON ENTROPY
#  Subset a dataframe with the key information
ML_otu_shannon_stats <- filter(otu_alphadiv, 
                               measure == "Shannon_Entropy" & 
                                 project == "Muskegon_Lake" & 
                                 fraction %in% c("WholePart", "WholeFree") & 
                                 year == "2015")

# Can fractional production be predicted by simpsevenness? 
free_ML_otu_shannon_stats <- filter(ML_otu_shannon_stats, fraction == "WholeFree")
freeprod_ML_otu_shannon <- lm(frac_bacprod ~ mean, data = free_ML_otu_shannon_stats)
freeprod_ML_otu_shannon
summary(freeprod_ML_otu_shannon)

part_ML_abs_shannon_stats <- filter(ML_otu_shannon_stats, fraction == "WholePart")
partprod_MLotu_shannon <- lm(frac_bacprod ~ mean, data = part_ML_abs_shannon_stats)
partprod_MLotu_shannon
summary(partprod_MLotu_shannon)

# Plot 
otu_shannon_vegan <- ggplot(ML_otu_shannon_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ggtitle("OTU: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  #scale_x_continuous(expand = c(0,0), limits=c(3, 6)) + 
  #scale_y_continuous(limits = c(0,70),expand = c(0,0)) + 
  ylab("Production (μgC/L/hr)") + xlab("Shannon Entropy") +
  #geom_smooth(data=subset(ML_otu_shannon_stats, fraction == "Free"), method='lm') + 
  geom_smooth(data=subset(ML_otu_shannon_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 4.5, y=45, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_shannon)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_shannon)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 5.35, y=5, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_shannon)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_shannon)$coefficients[,4][2]), digits = 4))); 





######################################################### INVERSE SIMPSON
#  Subset a dataframe with the key information
ML_otu_invsimps_stats <- filter(otu_alphadiv, 
                                measure == "Inverse_Simpson" & 
                                  project == "Muskegon_Lake" & 
                                  fraction %in% c("WholePart", "WholeFree") & 
                                  year == "2015")

# Can fractional production be predicted by invsimpsness? 
free_ML_otu_invsimps_stats <- filter(ML_otu_invsimps_stats, fraction == "WholeFree")
freeprod_ML_otu_invsimps <- lm(frac_bacprod ~ mean, data = free_ML_otu_invsimps_stats)
freeprod_ML_otu_invsimps
summary(freeprod_ML_otu_invsimps)

part_ML_abs_invsimps_stats <- filter(ML_otu_invsimps_stats, fraction == "WholePart")
partprod_MLotu_invsimps <- lm(frac_bacprod ~ mean, data = part_ML_abs_invsimps_stats)
partprod_MLotu_invsimps
summary(partprod_MLotu_invsimps)

# Plot Simpson's Evenness
otu_invsimps_vegan <- ggplot(ML_otu_invsimps_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ggtitle("OTU: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  scale_x_continuous(limits = c(0,100), expand = c(0,0)) + 
  #scale_y_continuous(limits = c(0,70),expand = c(0,0)) + 
  ylab("Production (μgC/L/hr)") + xlab("Inverse Simpson") +
  #geom_smooth(data=subset(ML_otu_invsimps_stats, fraction == "Free"), method='lm') + 
  geom_smooth(data=subset(ML_otu_invsimps_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.85,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 58, y=35, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_invsimps)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_invsimps)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 63, y=5, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_invsimps)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_invsimps)$coefficients[,4][2]), digits = 4))); 




######################################################### SIMPSON'S EVENNESS
#  Subset a dataframe with the key information
ML_otu_simpseven_stats <- filter(otu_alphadiv, 
                                 measure == "Simpsons_Evenness" & 
                                   project == "Muskegon_Lake" & 
                                   fraction %in% c("WholePart", "WholeFree") & 
                                   year == "2015")

# Can fractional production be predicted by simpsevenness? 
free_ML_otu_simpseven_stats <- filter(ML_otu_simpseven_stats, fraction == "WholeFree")
freeprod_ML_otu_simpseven <- lm(frac_bacprod ~ mean, data = free_ML_otu_simpseven_stats)
freeprod_ML_otu_simpseven
summary(freeprod_ML_otu_simpseven)

part_ML_abs_simpseven_stats <- filter(ML_otu_simpseven_stats, fraction == "WholePart")
partprod_MLotu_simpseven <- lm(frac_bacprod ~ mean, data = part_ML_abs_simpseven_stats)
partprod_MLotu_simpseven
summary(partprod_MLotu_simpseven)

# Plot 
otu_simpseven_vegan <- ggplot(ML_otu_simpseven_stats, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) +
  ggtitle("OTU: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("WholePart", "WholeFree")) +
  scale_x_continuous(expand = c(0,0), limits=c(0, 0.15)) + 
  #scale_y_continuous(limits = c(0,70),expand = c(0,0)) + 
  ylab("Production (μgC/L/hr)") + xlab("Simpson's Evenness") +
  #geom_smooth(data=subset(ML_otu_simpseven_stats, fraction == "Free"), method='lm') + 
  geom_smooth(data=subset(ML_otu_simpseven_stats, fraction == "WholePart"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 0.03, y=35, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_simpseven)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_simpseven)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 0.015, y=7, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_simpseven)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_simpseven)$coefficients[,4][2]), digits = 4))); 

otu_vegan <- plot_grid(otu_rich_vegan, otu_simpseven_vegan, otu_shannon_vegan, otu_invsimps_vegan,
                       labels = c("A", "B", "C", "D"), 
                       align = "h", nrow = 2, ncol = 2)
otu_vegan

ggsave("../Figures/veganRARE_otu_alpha_vs_prod.png", otu_vegan, dpi = 600, units = "in", width = 10, height = 8)
```


# Prefiltered Fraction Diversity-Production Analysis 
```{r PREFILTERED-otu-alphadiv, eval = TRUE, fig.width = 10, fig.height = 8, warning = FALSE}

###############################
ML_otu_rich_stats_prefilt <- filter(otu_alphadiv, measure == "Richness" & project == "Muskegon_Lake" & fraction %in% c("Particle", "Free") & year == "2015")

# Can fractional production be predicted by richness? 
freeprod_ML_otu_rich2 <- lm(frac_bacprod ~ mean, 
                            data = filter(ML_otu_rich_stats_prefilt, fraction == "Free"))
freeprod_ML_otu_rich2
summary(freeprod_ML_otu_rich2)


partprod_MLotu_rich2 <- lm(frac_bacprod ~ mean, 
                           data = filter(ML_otu_rich_stats_prefilt, fraction == "Particle"))
partprod_MLotu_rich2
summary(partprod_MLotu_rich2)

# Plot 
otu_rich_vegan_prefilt <- ggplot(ML_otu_rich_stats_prefilt, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ggtitle("Prefiltered: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("Particle", "Free")) +
  ylab("Production (μgC/L/hr)") + xlab("Observed Richness") +
  #geom_smooth(data=subset(ML_otu_rich_stats_prefilt, fraction == "Free"), method='lm') + 
  geom_smooth(data=subset(ML_otu_rich_stats_prefilt, fraction == "Particle"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 350, y=50, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_rich2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_rich2)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 500, y=50, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_rich2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_rich2)$coefficients[,4][2]), digits = 4)));



######################################################### SHANNON ENTROPY
#  Subset a dataframe with the key information
ML_otu_shannon_stats_prefilt <- filter(otu_alphadiv, 
                               measure == "Shannon_Entropy" & 
                                 project == "Muskegon_Lake" & 
                                 fraction %in% c("Particle", "Free") & 
                                 year == "2015")

# Can fractional production be predicted by simpsevenness? 

freeprod_ML_otu_shannon2 <- lm(frac_bacprod ~ mean, 
                              data = filter(ML_otu_shannon_stats_prefilt, fraction == "Free"))
freeprod_ML_otu_shannon2
summary(freeprod_ML_otu_shannon2)


partprod_MLotu_shannon2 <- lm(frac_bacprod ~ mean, 
                             data = filter(ML_otu_shannon_stats_prefilt, fraction == "Particle"))
partprod_MLotu_shannon2
summary(partprod_MLotu_shannon2)

# Plot 
otu_shannon_vegan_prefilt <- ggplot(ML_otu_shannon_stats_prefilt, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ggtitle("Prefiltered: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("Particle", "Free")) +
  #scale_x_continuous(expand = c(0,0), limits=c(3, 6)) + 
  #scale_y_continuous(limits = c(0,70),expand = c(0,0)) + 
  ylab("Production (μgC/L/hr)") + xlab("Shannon Entropy") +
  #geom_smooth(data=subset(ML_otu_shannon_stats_prefilt, fraction == "Free"), method='lm') + 
  geom_smooth(data=subset(ML_otu_shannon_stats_prefilt, fraction == "Particle"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 4.5, y=45, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_shannon2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_shannon2)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 4.9, y=55, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_shannon2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_shannon2)$coefficients[,4][2]), digits = 4))); 





######################################################### INVERSE SIMPSON
#  Subset a dataframe with the key information
ML_otu_invsimps_stats_prefilt <- filter(otu_alphadiv, 
                                measure == "Inverse_Simpson" & 
                                  project == "Muskegon_Lake" & 
                                  fraction %in% c("Particle", "Free") & 
                                  year == "2015")

# Can fractional production be predicted by invsimpsness? 
freeprod_ML_otu_invsimps2 <- lm(frac_bacprod ~ mean, 
                               data = filter(ML_otu_invsimps_stats_prefilt, fraction == "Free"))
freeprod_ML_otu_invsimps2
summary(freeprod_ML_otu_invsimps2)


partprod_MLotu_invsimps2 <- lm(frac_bacprod ~ mean, 
                               data = filter(ML_otu_invsimps_stats_prefilt, fraction == "Particle"))
partprod_MLotu_invsimps2
summary(partprod_MLotu_invsimps2)

# Plot Simpson's Evenness
otu_invsimps_vegan_prefilt <- ggplot(ML_otu_invsimps_stats_prefilt, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) + geom_errorbarh(aes(xmin = mean - sd, xmax = mean + sd)) + 
  ggtitle("Prefiltered: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("Particle", "Free")) +
  scale_x_continuous(limits = c(0,65), expand = c(0,0)) + 
  #scale_y_continuous(limits = c(0,70),expand = c(0,0)) + 
  ylab("Production (μgC/L/hr)") + xlab("Inverse Simpson") +
  #geom_smooth(data=subset(ML_otu_invsimps_stats, fraction == "Free"), method='lm') + 
  geom_smooth(data=subset(ML_otu_invsimps_stats_prefilt, fraction == "Particle"), method='lm') + 
  theme(legend.position=c(0.85,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 30, y=55, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_invsimps2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_invsimps2)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 21, y=3, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_invsimps2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_invsimps2)$coefficients[,4][2]), digits = 4))); 




######################################################### SIMPSON'S EVENNESS
#  Subset a dataframe with the key information
ML_otu_simpseven_stats_prefilt <- filter(otu_alphadiv, 
                                 measure == "Simpsons_Evenness" & 
                                   project == "Muskegon_Lake" & 
                                   fraction %in% c("Particle", "Free") & 
                                   year == "2015")

# Can fractional production be predicted by simpsevenness? 
freeprod_ML_otu_simpseven2 <- lm(frac_bacprod ~ mean, 
                                data = filter(ML_otu_simpseven_stats_prefilt, fraction == "Free"))
freeprod_ML_otu_simpseven2
summary(freeprod_ML_otu_simpseven2)


partprod_MLotu_simpseven2 <- lm(frac_bacprod ~ mean, 
                               data = filter(ML_otu_simpseven_stats_prefilt, fraction == "Particle"))
partprod_MLotu_simpseven2
summary(partprod_MLotu_simpseven2)

# Plot 
otu_simpseven_vegan_prefilt <- ggplot(ML_otu_simpseven_stats_prefilt, aes(x=mean, y=frac_bacprod, color = fraction)) + 
  geom_point(size = 3.5) +
  ggtitle("Prefiltered: Vegan") +
  scale_color_manual(values = c("firebrick3","cornflowerblue"), limits = c("Particle", "Free")) +

  ylab("Production (μgC/L/hr)") + xlab("Simpson's Evenness") +
  #geom_smooth(data=subset(ML_otu_simpseven_stats_prefilt, fraction == "Free"), method='lm') + 
  #geom_smooth(data=subset(ML_otu_simpseven_stats_prefilt, fraction == "Particle"), method='lm') + 
  theme(legend.position=c(0.15,0.9),        
        legend.title=element_blank()) +
  annotate("text", x = 0.09, y=45, color = "cornflowerblue", fontface = "bold",
           label = paste("R2 =", round(summary(freeprod_ML_otu_simpseven2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(freeprod_ML_otu_simpseven2)$coefficients[,4][2]), digits = 4))) + 
  annotate("text", x = 0.11, y=45, color = "firebrick3", fontface = "bold",
           label = paste("R2 =", round(summary(partprod_MLotu_simpseven2)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(partprod_MLotu_simpseven2)$coefficients[,4][2]), digits = 4))); 

otu_veganprefilt <- plot_grid(otu_rich_vegan_prefilt, otu_simpseven_vegan_prefilt,
                              otu_shannon_vegan_prefilt, otu_invsimps_vegan_prefilt,
                       labels = c("A", "B", "C", "D"), 
                       align = "h", nrow = 2, ncol = 2)
otu_veganprefilt
```




#  How linked are the diversities of the fractions?
## Free Living Sample Diversity Comparison  
```{r free_vs_wholefree_diversity, fig.height=8, fig.width=10, echo = TRUE}
otu_alphadiv <- select(otu_alphadiv, 1:10, 63:65)

# Plot diversity based only on fraction 
ggplot(filter(otu_alphadiv, fraction %in% c("Free","WholeFree")),
       aes(x = fraction, y = mean)) + 
  geom_jitter(aes(color = fraction), size = 3, width = 0.2) + 
  geom_boxplot(aes(fill = fraction), alpha =0.5) +
  scale_color_manual(values = fraction_colors) + 
  scale_fill_manual(values = fraction_colors) + 
  facet_wrap(~measure, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.07, 0.93), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold")) + 
  ggtitle("Rarefied at 2,895") 

# Plot diversity based on lakesite lumped together
ggplot(filter(otu_alphadiv, fraction %in% c("Free","WholeFree")),
       aes(x = lakesite, y = mean)) + 
  geom_jitter(aes(color = lakesite), size = 3, width = 0.2) + 
  geom_boxplot(aes(fill = lakesite), alpha =0.5) +
  scale_color_manual(values = lakesite_colors) + 
  scale_fill_manual(values = lakesite_colors) + 
  facet_wrap(~measure, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.07, 0.93), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold")) + 
  ggtitle("Rarefied at 2,895")


# Value of Diversity estimate on y-axis and site vs fraction on x-axis
ggplot(filter(otu_alphadiv, fraction %in% c("Free","WholeFree")),
       aes(x = lakesite, y = mean)) + 
  geom_jitter(aes(color = fraction), size = 3, position = position_dodge(width = 0.8)) + 
  geom_boxplot(aes(fill = fraction), alpha =0.5) +
  scale_color_manual(values = fraction_colors) + 
  scale_fill_manual(values = fraction_colors) + 
  facet_wrap(~measure, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.07, 0.93), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold")) + 
  ggtitle("Rarefied at 2,895")

###  ARE THERE DIFFERENCES BETWEEN FRACTION AND LAKESITE WITHIN THE FREE LIVING?
free_fraction_divs <- filter(otu_alphadiv, fraction %in% c("Free","WholeFree"))
#### Significant difference between Richness lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(free_fraction_divs, measure == "Richness")))
#### Significant difference between Simpsons_Evenness lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(free_fraction_divs, measure == "Simpsons_Evenness")))
#### Significant difference between Shannon_Entropy lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(free_fraction_divs, measure == "Shannon_Entropy")))
#### Significant difference between Inverse_Simpson lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(free_fraction_divs, measure == "Inverse_Simpson")))


# Prepare data frames for plotting diversity of wholefree vs free 
free_samps_alpha <- filter(otu_alphadiv, fraction == "Free" & year == "2015") %>% 
  dplyr::select(norep_filter_name, mean, measure, fraction) %>%
  spread(measure, mean) %>%
  # Make a new column called "measure" that includes the hill diversity metric
  gather("div_metric", "value", Richness, Shannon_Entropy, Inverse_Simpson, Simpsons_Evenness) %>%
  # Combine the diversity metric and fraction into one column called frac_div_metric
  unite("frac_div_metric", fraction, div_metric, sep = "_") %>%
  # Spread frac_div_metric into 4 columns for each hill div metric
  spread(frac_div_metric, value) %>%
  mutate(norep_water_name = paste(substr(norep_filter_name,1,4), substr(norep_filter_name, 6, 8), sep = "")) %>%
  dplyr::select(-norep_filter_name) 

# Prepare data frames for plotting diversity of wholefree vs free 
wholefree_samps_alpha <- filter(otu_alphadiv, fraction == "WholeFree" & year == "2015") %>% 
  dplyr::select(norep_filter_name, mean, measure, fraction, season, lakesite,year, limnion) %>%
  spread(measure, mean) %>%
  # Make a new column called "div_metric" that includes the hill diversity metric
  gather(div_metric, value, Richness, Shannon_Entropy, Inverse_Simpson, Simpsons_Evenness) %>%
  # Combine the diversity metric and fraction into one column called frac_div_metric
  unite(frac_div_metric, fraction, div_metric, sep = "_") %>%
  # Spread frac_div_metric into 4 columns for each hill div metric
  spread(frac_div_metric, value) %>%
  mutate(norep_water_name = paste(substr(norep_filter_name,1,4), substr(norep_filter_name, 6, 8), sep = "")) %>%
  dplyr::select(-norep_filter_name)

# join the data frames together 
FL_alpha <- left_join(free_samps_alpha, wholefree_samps_alpha, by = "norep_water_name")


# Is there a linear relationship between free and whole free D0 diversity?
lm_FL_Richness <- lm(Free_Richness ~ WholeFree_Richness, data = FL_alpha)
# Plot linear relationship between free and whole free D0 diversity
free_fraction_Richness <- ggplot(FL_alpha, aes(x = WholeFree_Richness, y = Free_Richness)) + 
  geom_point(size = 3) + geom_smooth(method = "lm") +
  #xlim(300, 2000) + ylim(300, 2000) + 
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 400, y=250, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_FL_Richness)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_FL_Richness)$coefficients[,4][2]), digits = 8))) + 
  ggtitle("Rarefied at 2,895")


# Is there a linear relationship between free and whole free D0_chao diversity?
lm_FL_simps_simpseven <- lm(Free_Simpsons_Evenness ~ WholeFree_Simpsons_Evenness, data = FL_alpha)
# Plot linear relationship between free and whole free D0 diversity
free_fraction_simpseven <- ggplot(FL_alpha, aes(x = WholeFree_Simpsons_Evenness, y = Free_Simpsons_Evenness)) + 
  geom_point(size = 3) + geom_smooth(method = "lm") +
  #xlim(300, 3600) + ylim(300, 3600) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 0.125, y=0.08, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_FL_simps_simpseven)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_FL_simps_simpseven)$coefficients[,4][2]), digits = 5))) + 
  ggtitle("Rarefied at 2,895")


# Is there a linear relationship between free and whole free D1 diversity?
lm_FL_shannon <- lm(Free_Shannon_Entropy ~ WholeFree_Shannon_Entropy, data = FL_alpha)
# Plot linear relationship between free and whole free D1 diversity
free_fraction_shannon <- ggplot(FL_alpha, aes(x = WholeFree_Shannon_Entropy, y = Free_Shannon_Entropy)) + 
  geom_point(size = 3) + geom_smooth(method = "lm") +
  #xlim(0, 175) + ylim(0,175) + 
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 4.25, y=3.5, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_FL_shannon)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_FL_shannon)$coefficients[,4][2]), digits = 12))) + 
  ggtitle("Rarefied at 2,895")

# Is there a linear relationship between free and whole free D2 diversity?
lm_FL_invsimps <- lm(Free_Inverse_Simpson ~ WholeFree_Inverse_Simpson, data = FL_alpha)
# Plot linear relationship between free and whole free D2 diversity
free_fraction_invsimps <- ggplot(FL_alpha, aes(x = WholeFree_Inverse_Simpson, y = Free_Inverse_Simpson)) + 
  geom_point(size = 3) + geom_smooth(method = "lm") +
  xlim(10, 55) + ylim(10, 55) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 35, y=17, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_FL_invsimps)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_FL_invsimps)$coefficients[,4][2]), digits = 8))) + 
  ggtitle("Rarefied at 2,895")
  
free_vs_freepart_div_plots <- plot_grid(free_fraction_Richness, free_fraction_simpseven, free_fraction_shannon, free_fraction_invsimps, 
          labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2)

free_vs_freepart_div_plots
#ggsave("../Figures/free_vs_wholefree_div_plots.png", plot = free_vs_freepart_div_plots, dpi = 600, width = 10, height = 8)

```

##  Particle-Associated Diversity Comparison 
```{r part_vs_wholepart_diversity, fig.height=8, fig.width=10, echo = TRUE}
# Plot diversity based only on fraction 
ggplot(filter(otu_alphadiv, fraction %in% c("Particle","WholePart")),
       aes(x = fraction, y = mean)) + 
  geom_jitter(aes(color = fraction), size = 3, width = 0.2) + 
  geom_boxplot(aes(fill = fraction), alpha =0.5) +
  scale_color_manual(values = fraction_colors) + 
  scale_fill_manual(values = fraction_colors) + 
  facet_wrap(~measure, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.07, 0.93), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold")) + 
  ggtitle("Rarefied at 2,895")

# Plot based on lakesite lumped together
ggplot(filter(otu_alphadiv, fraction %in% c("Particle","WholePart")),
       aes(x = lakesite, y = mean)) + 
  geom_jitter(aes(color = lakesite), size = 3, width = 0.2) + 
  geom_boxplot(aes(fill = lakesite), alpha =0.5) +
  scale_color_manual(values = lakesite_colors) + 
  scale_fill_manual(values = lakesite_colors) + 
  facet_wrap(~measure, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.07, 0.93), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold")) + 
  ggtitle("Rarefied at 2,895")


# Value of Diversity estimate on y-axis and site vs fraction on x-axis
ggplot(filter(otu_alphadiv, fraction %in% c("Particle","WholePart")),
       aes(x = lakesite, y = mean)) + 
  geom_jitter(aes(color = fraction), size = 3, position = position_dodge(width = 0.8)) + 
  geom_boxplot(aes(fill = fraction), alpha =0.5) +
  scale_color_manual(values = fraction_colors) + 
  scale_fill_manual(values = fraction_colors) + 
  facet_wrap(~measure, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.07, 0.93), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold")) + 
  ggtitle("Rarefied at 2,895")


###  ARE THERE DIFFERENCES BETWEEN FRACTION AND LAKESITE WITHIN PARTICLE ASSOCIATION?
part_fraction_divs <- filter(otu_alphadiv, fraction %in% c("Particle","WholePart"))

#### Significant difference between Richness lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(part_fraction_divs, measure == "Richness")))
#### Significant difference between Simpsons_Evenness lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(part_fraction_divs, measure == "Simpsons_Evenness")))
#### Significant difference between Shannon_Entropy lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(part_fraction_divs, measure == "Shannon_Entropy")))
#### Significant difference between Inverse_Simpson lakesite and fraction?
summary(aov(mean ~ fraction*lakesite,  
            data = filter(part_fraction_divs, measure == "Inverse_Simpson")))


# Prepare data frames for plotting diversity of wholepart vs particle 
part_samps_alpha <- filter(otu_alphadiv, fraction == "Particle" & year == "2015") %>% # & norep_filter_name != "MBRHJ515"
  dplyr::select(norep_filter_name, mean, measure, fraction) %>%
  spread(measure, mean) %>%
  gather(measure, value, Richness, Shannon_Entropy, Inverse_Simpson, Simpsons_Evenness) %>%
  unite("frac_div_metric", fraction, measure, sep = "_") %>%
  spread(frac_div_metric, value) %>%
  mutate(norep_water_name = paste(substr(norep_filter_name,1,4), substr(norep_filter_name, 6, 8), sep = "")) %>%
  select(-norep_filter_name)

# Prepare data frames for plotting diversity of wholepart vs particle 
wholepart_samps_alpha <- filter(otu_alphadiv, fraction == "WholePart" & year == "2015") %>% # & norep_filter_name != "MBRHJ515"
  dplyr::select(norep_filter_name, mean, measure, fraction, season, lakesite,year, limnion) %>%
  spread(measure, mean) %>%
  gather(measure, value, Richness, Shannon_Entropy, Inverse_Simpson, Simpsons_Evenness) %>%
  unite(frac_div_metric, fraction, measure, sep = "_") %>%
  spread(frac_div_metric, value) %>%
  mutate(norep_water_name = paste(substr(norep_filter_name,1,4), substr(norep_filter_name, 6, 8), sep = "")) %>%
  select(-norep_filter_name)

# Combine the data frames into one
PA_alpha <- left_join(part_samps_alpha, wholepart_samps_alpha, by = "norep_water_name")

# Is there a linear relationship between particle and WholePart D0 diversity?
lm_PA_richness <- lm(Particle_Richness ~ WholePart_Richness, data = PA_alpha)
# Plot the linear relationship between particle and WholePart D0 diversity
PA_fraction_richness <- ggplot(PA_alpha, aes(x = WholePart_Richness, y = Particle_Richness)) + 
  geom_point(size = 3) + geom_smooth(method = "lm", se = TRUE) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 650, y=350, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_PA_richness)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_PA_richness)$coefficients[,4][2]), digits = 8))) + 
  ggtitle("Rarefied at 2,895")


# Is there a linear relationship between particle and WholePart D0_chao diversity?
lm_PA_simpseven <- lm(Particle_Simpsons_Evenness ~ WholePart_Simpsons_Evenness, 
                                           data = filter(PA_alpha, norep_water_name != "MBRH515"))
# Plot the linear relationship between particle and WholePart D0_chao diversity
PA_fraction_simpseven <- ggplot(filter(PA_alpha, norep_water_name != "MBRH515"),
                                aes(x = WholePart_Simpsons_Evenness, y = Particle_Simpsons_Evenness)) + 
  geom_point(size = 3) + geom_smooth(method = "lm", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 0.12, y=0.05, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_PA_simpseven)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_PA_simpseven)$coefficients[,4][2]), digits = 4))) + 
  ggtitle("Rarefied at 2,895 \n (removed MBRH515)")


# Is there a linear relationship between particle and WholePart D1 diversity?
lm_PA_shannon <- lm(Particle_Shannon_Entropy ~ WholePart_Shannon_Entropy, data = PA_alpha)
# Plot the linear relationship between particle and WholePart D1 diversity
PA_fraction_shannon <- ggplot(PA_alpha, aes(x = WholePart_Shannon_Entropy, y = Particle_Shannon_Entropy)) + 
  geom_point(size = 3) + geom_smooth(method = "lm") + 
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", y = 4.5, x=5.5, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_PA_shannon)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_PA_shannon)$coefficients[,4][2]), digits = 5))) + 
  ggtitle("Rarefied at 2,895")

# Is there a linear relationship between particle and WholePart D2 diversity?
lm_PA_invsimps <- lm(Particle_Inverse_Simpson ~ WholePart_Inverse_Simpson, 
                     data = filter(PA_alpha, WholePart_Inverse_Simpson < 219))
# Plot the linear relationship between particle and WholePart D2 diversity
PA_fraction_invsimps <- ggplot(PA_alpha, aes(x = WholePart_Inverse_Simpson, y = Particle_Inverse_Simpson)) + 
  geom_point(size = 3) + geom_smooth(method = "lm") +
  xlim(0, 100) + ylim(0, 100) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 75, y=25, color = "black", fontface = "bold",
           label = paste("R2 =", round(summary(lm_PA_invsimps)$adj.r.squared, digits = 4), "\n", 
                         "p =", round(unname(summary(lm_PA_invsimps)$coefficients[,4][2]), digits = 5))) + 
  ggtitle("Rarefied at 2,895 \n(removed MBRH515)")

  
part_vs_wholepart_div_plots <- plot_grid(PA_fraction_richness, PA_fraction_simpseven,
                                         PA_fraction_shannon, PA_fraction_invsimps, 
          labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2)

part_vs_wholepart_div_plots
#ggsave("../Figures/part_vs_wholepart_div_plots.png", plot = part_vs_wholepart_div_plots, dpi = 600, width = 10, height = 8)
```


# Is there a significant relationship between sequencing depth and diversity?
```{r div_vs_seqdepth, fig.width=10, fig.height = 4, echo = TRUE}
## Is there a significant relationship between sequencing depth and D0?
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Richness" & fraction == "Free"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Richness" & fraction == "WholeFree"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Richness" & fraction == "Particle"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Richness" & fraction == "WholePart"))) # NS

# Plot Sequencing depth vs Richness
ggplot(filter(otu_alphadiv, measure == "Richness"), 
       aes(x = Sample_TotalSeqs, y = mean)) + 
  geom_point(size = 3) + 
  facet_grid(.~fraction, scale = "free_x") + 
  ggtitle("Observed Richness") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.1, 0.9), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold"))

## Is there a significant relationship between sequencing depth and Simpsons_Evenness?
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Simpsons_Evenness" & fraction == "Free"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Simpsons_Evenness" & fraction == "WholeFree"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Simpsons_Evenness" & fraction == "Particle"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Simpsons_Evenness" & fraction == "WholePart"))) # NS

# Plot the relationship between Simpsons_Evenness and sequencing depth 
ggplot(filter(otu_alphadiv, measure == "Simpsons_Evenness"), 
       aes(x = Sample_TotalSeqs, y = mean)) + 
  geom_point(size = 3) + 
  facet_grid(.~fraction, scale = "free_x") + 
  ggtitle("Simpsons_Evenness") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.1, 0.9), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold"))

## Is there a significant relationship between sequencing depth and Shannon_Entropy?
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Shannon_Entropy" & fraction == "Free"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Shannon_Entropy" & fraction == "WholeFree"))) # NS 
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Shannon_Entropy" & fraction == "Particle"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Shannon_Entropy" & fraction == "WholePart"))) # NS

# Plot the relationship between D1 and sequencing depth 
ggplot(filter(otu_alphadiv, measure == "Shannon_Entropy"), 
       aes(x = Sample_TotalSeqs, y = mean)) + 
  geom_point(size = 3) + 
  facet_grid(.~fraction, scale = "free_x") + 
  ggtitle("Shannon Entropy") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.1, 0.9), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold"))

## Is there a significant relationship between sequencing depth and D2?
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Inverse_Simpson" & fraction == "Free"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Inverse_Simpson" & fraction == "WholeFree"))) # NS 
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Inverse_Simpson" & fraction == "Particle"))) # NS
#summary(lm(mean ~ Sample_TotalSeqs,data = filter(otu_alphadiv, measure == "Inverse_Simpson" & fraction == "WholePart"))) # NS

# Plot the relationship between D2 and sequencing depth 
ggplot(filter(otu_alphadiv, measure == "Inverse_Simpson"), 
       aes(x = Sample_TotalSeqs, y = mean)) + 
  geom_point(size = 3) + 
  facet_grid(.~fraction, scale = "free_x") + 
  ggtitle("Inverse Simpson") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = c(0.1, 0.9), 
        strip.background = element_rect(fill = NA), 
        strip.text.x = element_text(face = "bold"))

```

# Does DNA extraction concentration influcence diversity?
```{r dnaconc_vs_diversity, fig.width=10, fig.height = 4, echo = TRUE}
### Prepare the data frame 
DNA_comparison <- nosed_meta_data %>%
  dplyr::select(norep_filter_name, dnaconcrep1, dnaconcrep2, cells_per_uL, lakesite) %>%
  mutate(dnaconcrep1 = as.numeric(dnaconcrep1),
         dnaconcrep2 = as.numeric(dnaconcrep2),
         mean_dna = (dnaconcrep1+dnaconcrep2)/2)

# Calculate the linear models 
lm_dna_conc <- lm(dnaconcrep1 ~ dnaconcrep2,  data = DNA_comparison)
lm_cells_vs_meanDNA <- lm(mean_dna ~ cells_per_uL,  data = DNA_comparison)

# Plot the trends 
plot_grid(ggplot(DNA_comparison, aes(x = dnaconcrep1, y = dnaconcrep2)) + # First plot 
            geom_point(size = 3) + geom_smooth(method = "lm") +
            annotate("text",  x = 17, y = 3,
              color = "black", fontface = "bold",
               label = paste("R2 =", round(summary(lm_dna_conc)$adj.r.squared, digits = 4), "\n", 
                             "p =", round(unname(summary(lm_dna_conc)$coefficients[,4][2]), digits = 8)))+
            ggtitle("DNA Concentration Between Replicates"), 
          ggplot(DNA_comparison, aes(x = mean_dna, y = cells_per_uL)) + # Second plot 
            geom_point(size = 3) + geom_smooth(method = "lm") +
            annotate("text",  x = 12, y = 4800,
              color = "black", fontface = "bold",
               label = paste("R2 =", round(summary(lm_cells_vs_meanDNA)$adj.r.squared, digits = 4), "\n", 
                             "p =", round(unname(summary(lm_cells_vs_meanDNA)$coefficients[,4][2]), digits = 4)))+
            ggtitle("Mean DNA Concentration vs Cell Numbers"),
          labels = c("A", "B"), ncol = 2)
```


